---
title: "extra code"
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
---

# Exploratory Data Analyses

## Log Cytokine Conc ~ Variables
```{r, warning=FALSE}
# Make data long format
datalong <-  melt(data, id = c("Sample", "year","sex", "location", "molt.stage", 
                               "iav", "iavser", "bodcond", "bc_bin", "stage"), 
                  variable.name="cytokine", value.name="conc") %>% 
  mutate(conc=as.numeric(conc),
         logconc=log(conc+0.001))

# Plot~
year <- ggplot(datalong, aes(x=year, y=logconc, group=year)) +
  geom_jitter(width=0.15) +
  labs(x="Year", y="Cytokine Conc (pg/mL)") +
  facet_wrap(~cytokine, scales="free_y", ncol=5) +
  theme(axis.text.x=element_text(angle=45, hjust=1))
year

sex <- datalong %>%
  na.omit(datalong$sex) %>%
  ggplot(., aes(x=sex, y=logconc, group=sex)) +
  geom_jitter(width=0.2) +
  labs(x="Sex", y="Log Cytokine Concentration") +
  facet_wrap(~cytokine, scales="free_y", ncol=5)
sex

molt <- datalong %>%
  na.omit(datalong$molt.stage) %>%
  ggplot(., aes(x=molt.stage, y=logconc, group=molt.stage)) +
  geom_jitter(width=0.2) +
  labs(x="Molt Stage", y="Log Cytokine Concentration") +
  facet_wrap(~cytokine, scales="free_y", ncol=5)
molt

loc <- ggplot(datalong, aes(x=location, y=logconc, group=location)) +
  geom_jitter(width=0.2) +
  labs(x="Location", y="Log Cytokine Concentration") +
  facet_wrap(~cytokine, scales="free_y", ncol=5)
loc

iav <- ggplot(datalong, aes(x=iav, y=logconc, group=iav)) +
  geom_jitter(width=0.2) +
  labs(x="IAV", y="Log Cytokine Concentration") +
  facet_wrap(~cytokine, scales="free_y", ncol=5) 
iav

bc <- ggplot(datalong, aes(x=bodcond, y=logconc)) +
  geom_jitter(width=0.2) +
  labs(x="Body Condition", y="Log Cytokine Concentration")+
  facet_wrap(~cytokine, scales="free_y", ncol=5)
bc

bc_bin <- ggplot(datalong, aes(x=bc_bin, y=logconc)) +
  geom_boxplot() +
  labs(x="Body Condition Binned", y="Log Cytokine Concentration")+
  facet_wrap(~cytokine, scales="free_y", ncol=5)
bc_bin

```

## Binary Cytokine ~ Variables
```{r}
databin_plot <- databin[,-1]

# Influenza Virus Infection
plots <- vector('list', length=13)
for (i in 1:13) {
  tab <- 
    data.frame(table(databin_plot$iav, databin_plot[,i])) %>% 
    rename("Cytokine" = "Var2")
  plots[[i]] <- ggplot(data=tab, aes(x=Var1, y=Freq, fill=Cytokine)) +
    geom_col(position="fill") +
    labs(x=NULL, y=NULL, title=paste(colnames(databin_plot)[i]))
}
  
grid.arrange(arrangeGrob(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], plots[[11]], plots[[12]], plots[[13]]), left=text_grob("Proportion", rot=90), bottom=text_grob("IAV"))

# Year
plots <- vector('list', length=13)
for (i in 1:13) {
  tab <- 
    data.frame(table(databin_plot$year, databin_plot[,i])) %>% 
    rename("Cytokine" = "Var2")
  plots[[i]] <- ggplot(data=tab, aes(x=Var1, y=Freq, fill=Cytokine)) +
    geom_col(position="fill") +
    labs(x=NULL, y=NULL, title=paste(colnames(databin_plot)[i])) +
    theme(axis.text.x=element_text(angle=45, hjust=1))

}
  
grid.arrange(arrangeGrob(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], plots[[11]], plots[[12]], plots[[13]]), left=text_grob("Proportion", rot=90), bottom=text_grob("Year"))

# Location
plots <- vector('list', length=13)
for (i in 1:13) {
  tab <- 
    data.frame(table(databin_plot$location, databin_plot[,i])) %>% 
    rename("Cytokine" = "Var2")
  plots[[i]] <- ggplot(data=tab, aes(x=Var1, y=Freq, fill=Cytokine)) +
    geom_col(position="fill") +
    labs(x=NULL, y=NULL, title=paste(colnames(databin_plot)[i])) +
    theme(axis.text.x=element_text(angle=45, hjust=1))

}
  
grid.arrange(arrangeGrob(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], plots[[11]], plots[[12]], plots[[13]]), left=text_grob("Proportion", rot=90), bottom=text_grob("Location"))

# Sex
plots <- vector('list', length=13)
for (i in 1:13) {
  tab <- 
    data.frame(table(databin_plot$sex, databin_plot[,i])) %>% 
    rename("Cytokine" = "Var2")
  plots[[i]] <- ggplot(data=tab, aes(x=Var1, y=Freq, fill=Cytokine)) +
    geom_col(position="fill") +
    labs(x=NULL, y=NULL, title=paste(colnames(databin_plot)[i]))
}
  
grid.arrange(arrangeGrob(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], plots[[11]], plots[[12]], plots[[13]]), left=text_grob("Proportion", rot=90), bottom=text_grob("Sex"))


# Molt Stage
plots <- vector('list', length=13)
for (i in 1:13) {
  tab <- 
    data.frame(table(databin_plot$molt.stage, databin_plot[,i])) %>% 
    rename("Cytokine" = "Var2")
  plots[[i]] <- ggplot(data=tab, aes(x=Var1, y=Freq, fill=Cytokine)) +
    geom_col(position="fill") +
    labs(x=NULL, y=NULL, title=paste(colnames(databin_plot)[i]))
}
  
grid.arrange(arrangeGrob(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], plots[[11]], plots[[12]], plots[[13]]), left=text_grob("Proportion", rot=90), bottom=text_grob("Molt Stage"))


# Body Condition
databin_long <- melt(databin, id=c("Sample", "year", "sex", "location", "molt.stage", "iav", 
                                   "iavser", "bodcond", "bc_bin"),
                     variable.name="cytokine", value.name="presence") %>% 
  mutate(presence=as.factor(presence))

bc_cytopres <- ggplot(data=databin_long, aes(x=presence, y=bodcond, group=presence)) +
  geom_boxplot() +
  labs(x="Cytokine Presence/Absence", y="Body Condition") +
  facet_wrap(~cytokine, scales="free_y", ncol=5)
bc_cytopres


# Body Condition Binned
plots <- vector('list', length=13)
for (i in 1:13) {
  tab <- 
    data.frame(table(databin_plot$bc_bin, databin_plot[,i])) %>% 
    rename("Cytokine" = "Var2")
  plots[[i]] <- ggplot(data=tab, aes(x=Var1, y=Freq, fill=Cytokine)) +
    geom_col(position="fill") +
    labs(x=NULL, y=NULL, title=paste(colnames(databin_plot)[i]))
}
  
grid.arrange(arrangeGrob(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], plots[[11]], plots[[12]], plots[[13]]), left=text_grob("Proportion", rot=90), bottom=text_grob("Body Condition"))

```

## Th1/Th2 Responses ~ Variables
  - Th1: IL2, IL15, IL18, IFNg, IP-10, TNFa 
  - Th2: IL6, IL10 
  - other: GM.CSF, IL7, IL8, KC.like, MCP1 
```{r, warning=FALSE}
# Configure data
data_th <-
  data%>%
  mutate(th1=IL.2 + IL.15 + IL.18 + IFNg + IP.10 + TNFa, 
         th2=IL.6 + IL.10,
         other=GM.CSF + IL.7 + IL.8 + KC.like + MCP.1)%>%
  select(Sample, iav, iavser, location, year, sex, molt.stage, 
         th1, th2, other, bc_bin) %>%
  melt(id.vars=c("Sample","iav","iavser","location","year","sex","molt.stage", "bc_bin"), variable.name="cytokine", value.name="concentration") %>%
  mutate(logconc = log(concentration))

# IAV+ vs IAV-
th1vth2_sex <- 
  data_th %>%
  na.omit(data_th$sex) %>%
  ggplot(., aes(x=sex, y=logconc, fill=cytokine)) +
  geom_boxplot()

th1vth2_location <- 
  data_th %>%
  na.omit(data_th$location) %>%
  ggplot(., aes(x=location, y=logconc, fill=cytokine)) +
  geom_boxplot()

th1vth2_molt <- 
  data_th %>%
  na.omit(data_th$molt.stage) %>%
  ggplot(., aes(x=molt.stage, y=logconc, fill=cytokine)) +
  geom_boxplot()

th1vth2_year <- 
  ggplot(data_th, aes(x=year, y=logconc, fill=cytokine)) +
  geom_boxplot()

th1vth2_bodcond <-
 ggplot(data_th, aes(x=bc_bin, y=logconc, fill=cytokine)) +
  geom_boxplot()

 grid.arrange(arrangeGrob(th1vth2_bodcond, th1vth2_sex, ncol=2), 
             arrangeGrob(th1vth2_location, th1vth2_molt, ncol=2),
             arrangeGrob(th1vth2_year, ncol=1))

th1vth2_iav <-
  ggplot(data=data_th, aes(x=iav, y=logconc, fill=cytokine)) +
  geom_boxplot()
th1vth2_iav
```


## Test assumption of normality for cytokine concentrations
  - Shapiro-Wilk 
  - all p < 0.05 - all data not normally distributed

```{r}
# Configure data
data_norm <- 
  data %>%
  select(2:14) %>% 
  select_if(function(x) sum(x>0) > 11)

# Test
norm_results <- data.frame(matrix(ncol=0, nrow=9)) %>%
  mutate(cyto=colnames(data_norm)[1:9],
         W=NA, 
         pvalue=NA)

for (i in 1:9) {
  test <- shapiro.test(data_norm[,i])
  norm_results[i,2] <- round(test$statistic, digits=3)
  norm_results[i,3] <- test$p.value
}

norm_results
```




## Hurdle Model - cytokines ~ factors

### Cyto (p/a) ~ factors
```{r, warning=FALSE}
# Exclude IL-18 - present in all seals, no absence
data_pa <- 
  databin %>%
  column_to_rownames(var="Sample") %>%
  select(!IL.18)

pa_results <- data.frame(matrix(ncol=0, nrow=12)) %>%
  mutate(cyto=colnames(data_pa)[1:12],
         sex=NA,
         location=NA,
         molt.stage=NA, 
         bodcond=NA,
         bodcond_ord=NA)

# Sex
for (i in 1:12) {
tab <- table(data_pa$sex, data_pa[,i])
pa_results[i,2] <- 
  tryCatch({
  round(chisq.test(tab)$p.value, digits=2)
}, warning = function(w) {
  round(fisher.test(tab)$p.value, digits=2)
})
}

# Location
for (i in 1:12) {
tab <- table(data_pa$location, data_pa[,i])
pa_results[i,3] <- 
  tryCatch({
  round(chisq.test(tab)$p.value, digits=2)
}, warning = function(w) {
  round(fisher.test(tab)$p.value, digits=2)
})
}

# Molt Stage
for (i in 1:12) {
tab <- table(data_pa$molt.stage, data_pa[,i])
pa_results[i,4] <- 
  tryCatch({
  round(chisq.test(tab)$p.value, digits=2)
}, warning = function(w) {
  round(fisher.test(tab)$p.value, digits=2)
})
}

# Body condition
for (i in 1:12) {
tab <- table(data_pa$bc_bin, data_pa[,i])
pa_results[i,5] <- 
  tryCatch({
  round(chisq.test(tab)$p.value, digits=2)
}, warning = function(w) {
  round(fisher.test(tab)$p.value, digits=2)
})
}

pa_results

# Body Condition as ordinal variable (cyto p/a ~ bc)
data_pa_ord <- 
  data_pa %>% 
  mutate(bc_ord=ifelse(bc_bin=="lower", 1, 
                   ifelse(bc_bin=="middle", 2, 
                          ifelse(bc_bin=="upper", 3, NA))))
for (i in 1:12) {
  lbl_data <- table(data_pa_ord$bc_ord, data_pa_ord[,i])
  pa_results[i,6] <- round(pvalue(lbl_test(lbl_data)), digits=2)
}
pa_results


# Post-hoc pairwise tests for molt stage (most interesting)
pairwise_fisher_test(table(data_pa$molt.stage, data_pa$GM.CSF)) 
pairwise_fisher_test(table(data_pa$molt.stage, data_pa$IL.8)) 
pairwise_fisher_test(table(data_pa$molt.stage, data_pa$IL.15)) 
pairwise_fisher_test(table(data_pa$molt.stage, data_pa$KC.like)) 

# Plot! 
gmcsf_molt_plot <-
  data_pa %>% 
  mutate(GM.CSF=factor(GM.CSF, levels=c("1", "0"))) %>% 
  ggplot(data=.) +
  geom_mosaic(aes(x=product(GM.CSF,molt.stage), fill=GM.CSF)) +
  scale_fill_manual(values=c("0"="#b4b4b4", "1"="#7eaaac")) +
  labs(x="Molt Stage", y="Cytokine Presence/Absence", fill="GM.CSF") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
il8_molt_plot <-
  data_pa %>% 
  mutate(IL.8=factor(IL.8, levels=c("1", "0"))) %>% 
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.8,molt.stage), fill=IL.8)) +
  scale_fill_manual(values=c("0"="#b4b4b4", "1"="#7eaaac")) +
  labs(x="Molt Stage", y="Cytokine Presence/Absence", fill="IL.8") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

il15_molt_plot <-
  data_pa %>% 
  mutate(IL.15=factor(IL.15, levels=c("1", "0"))) %>% 
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.15,molt.stage), fill=IL.15)) +
  scale_fill_manual(values=c("0"="#b4b4b4", "1"="#7eaaac")) +
  labs(x="Molt Stage", y="Cytokine Presence/Absence", fill="IL.15") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
kclike_molt_plot <-
  data_pa %>% 
  mutate(KC.like=factor(KC.like, levels=c("1", "0"))) %>% 
  ggplot(data=.) +
  geom_mosaic(aes(x=product(KC.like,molt.stage), fill=KC.like)) +
  scale_fill_manual(values=c("0"="#b4b4b4", "1"="#7eaaac")) +
  labs(x="Molt Stage", y="Cytokine Presence/Absence", fill="KC.like") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
  
grid.arrange(gmcsf_molt_plot, il8_molt_plot, il15_molt_plot, kclike_molt_plot, ncol=2)


# Post-hoc pairwise tests for body condition
pairwise_fisher_test(table(data_pa$bc_bin, data_pa$IL.6)) 
pairwise_fisher_test(table(data_pa$bc_bin, data_pa$IP.10)) 

il6_bc_plot <-
  data_pa %>% 
  mutate(IL.6=factor(IL.6, levels=c("1", "0"))) %>% 
  na.omit(IL.6) %>% 
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.6,bc_bin), fill=IL.6)) +
  scale_fill_manual(values=c("0"="#b4b4b4", "1"="#7eaaac")) +
  labs(x="Molt Stage", y="Cytokine Presence/Absence", fill="IL.6") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
il6_bc_plot
```

### Cyto (nonzero) ~ factors
  - non-parametric anova 
  - excluding GM-CSF, IP-10, MCP-1, TNFa (3 or less detections) 
```{r, warning=FALSE}
conc_var_results <- 
  data.frame(matrix(nrow=5, ncol=0), 
             row.names=c("year", "sex", "location", "molt.stage", "bc_bin")) %>%
  mutate(IFNg=NA,
         IL.2=NA,
         IL.6=NA,
         IL.7=NA,
         IL.8=NA,
         IL.10=NA,
         IL.15=NA,
         IL.18=NA,
         KC.like=NA) 

  subdata <- 
  data %>%
  select(IFNg, IL.2, IL.6,  IL.7, IL.8, IL.10, IL.15, IL.18, KC.like,
         year, sex, location, molt.stage, bc_bin)

for (i in 1:9){
  nonzero <- subdata %>% filter(subdata[,i]>0)
  conc_var_results[1,i] <- round(kruskal.test(nonzero[,i] ~ nonzero$year)$p.value, digits=2)
  conc_var_results[2,i] <- round(wilcox.test(nonzero[,i] ~ nonzero$sex)$p.value, digits=2)
  conc_var_results[3,i] <- round(kruskal.test(nonzero[,i] ~ nonzero$location)$p.value, digits=2)
  conc_var_results[4,i] <- round(kruskal.test(nonzero[,i] ~ nonzero$molt.stage)$p.value, digits=2)
  conc_var_results[5,i] <- round(kruskal.test(nonzero[,i] ~ nonzero$bc_bin)$p.value, digits=2)
}
conc_var_results

# Plot molt
il18_molt_box <- ggplot(data=subdata, aes(x=molt.stage, y=IL.18)) + 
  geom_boxplot() 

# Dunn's test for multiple comparisons (holm=default)
DunnTest(subdata$IL.18~molt.stage, subdata, method="holm")

```
