---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Tidy Cytokine Data Metadata

## Libraries
```{r, message=FALSE, warning=FALSE} 
library(tidyverse)
```

## Raw file import & cleaning
```{r, message=FALSE, warning=FALSE}
cyto1 <- read.csv("Input Files/Grey seal cytokine 200_300_800_900.csv")
cyto1 <- cyto1[-c(1:6, 8:18, 37:56),-c(1:2)]
colnames(cyto1) <- cyto1[1,]
colnames(cyto1)[1] <- "Sample"
cyto1 <- cyto1[-1,]

cyto2 <- read.csv("Input Files/Results #1349 plate 1 grey seal millipore canine kit.csv")
cyto2 <- cyto2[, -c(2:3, 5:6, 8:10)]
cyto2 <- reshape(cyto2, idvar="Description", timevar="Analyte", direction="wide")
colnames(cyto2) <- colnames(cyto1)
cyto2 <- cyto2[-c(1:3),]
# these samples quantified twice. kept samples from first run (ML - cyto1)
cyto2 <- filter(cyto2, !Sample=="853" 
                     & !Sample=="869"
                     & !Sample=="896"
                     & !Sample=="900") 

cyto3 <- read.csv("Input Files/Results #1350 plate 2 grey seal millipore canine kit.csv")
cyto3 <- cyto3[-1,-c(2:3, 5:6, 8:10)]
cyto3 <- reshape(cyto3, idvar="Description", timevar="Analyte", direction="wide")
colnames(cyto3) <- colnames(cyto1)
cyto3 <- cyto3[-c(1:3),]

cyto4 <- read.csv("Input Files/Results #1359 4-13-23 2023 Monomoy_Muskeget_GP cytokines.csv")
cyto4 <- cyto4[-c(1:12), -c(1:2,17)]
colnames(cyto4) <- colnames(cyto1)
for (i in 1:nrow(cyto4)) {
  str <- cyto4[i,1]
  new_str <- gsub('Hg ','',str)
  cyto4[i,1] <- new_str
}
```

## Merge cytokine data & tidy
```{r}
cyto <- 
  bind_rows(cyto1, cyto2, cyto3, cyto4) 

# human error while typing in sample names to computer 1505 -> 1507
cyto$Sample[74] <- "1507" 

# Remove () from cytokine names
cytokines <- as.data.frame(colnames(cyto[-1]))
for (i in 1:13) {
  str <- cytokines[i,]
  new_str <- gsub("\\s*\\([^\\)]+\\)","",str)
  cytokines[i,] <- new_str
}
colnames(cyto)[2:14] <- cytokines$`colnames(cyto[-1])`

# Change OOR < to 0 & remove *
for (i in 1:nrow(cyto)) {
  for (j in 1:ncol(cyto)) {
    str <- cyto[i,j]
    new_str <- (gsub("\\*","",str))
    cyto[i,j] <- new_str
    {if (cyto[i,j] == "OOR <") {
      cyto[i,j] <- 0
    }
      else{next}
    }
  }
}

# Set cyto conc to as numeric 
cyto[,2:14] <- sapply(cyto[,2:14], as.numeric)

# Order cytokines alphabetically 
cyto <- 
  cyto %>% 
  select(Sample, `GM-CSF`, IFNg, `IL-2`, `IL-6`, `IL-7`, `IL-8`, `IL-10`, 
          `IL-15`, `IL-18`, `IP-10`, `KC-like`, `MCP-1`, TNFa)

# Write out data for all cytokines
write.csv(cyto, "Output Files/cleaned_cytokine_all.csv", row.names=FALSE)

# Remove cytokines with <10% (11) detections
cyto_10 <- 
  cyto %>% 
  select_if(function(x) sum(x>0) > 11) %>% 
  mutate(Sample = cyto$Sample)
  

# Write out reduced cytokine data
write.csv(cyto_10, file="Output Files/cleaned_cytokine.csv", row.names=FALSE)
```

## Create Cytokine (0/1) Dataset
```{r}
databin <- cyto
for (i in 1:nrow(databin)) {
  for (j in 2:14) {
  databin[i,j] <- ifelse(databin[i,j] > 0, 1, 0)
  }
}

# Write out binary cytokine data for all
write.csv(databin, "Output Files/cleaned_cytokine_bin_all.csv", row.names=FALSE)


# Remove cytokines with <10% (11) detections
databin_10 <- 
  databin %>% 
  select_if(function(x) sum(x>0) > 11) 


# Write out binary cytokine dataset
write.csv(databin_10, file="Output Files/cleaned_cytokine_bin.csv", row.names=FALSE)
```

## Tidy Metadata
```{r}
metadata <- 
  read.csv("Input Files/metadata.csv") %>% 
  mutate(bodcond = (ax.girth*100)/st.length,
         bc_bin = ifelse(bodcond < quantile(bodcond, probs=0.25, na.rm=TRUE), "Poor",
                         ifelse(bodcond > quantile(bodcond, probs=0.25, na.rm=TRUE) & 
                                  bodcond < quantile(bodcond, probs=0.75, na.rm=TRUE), "Average",
                          ifelse(bodcond > quantile(bodcond, probs=0.75, na.rm=TRUE), "Robust", NA))),
         stage = ifelse(iav=="neg" & iavser=="neg","control",
                        ifelse(iav=="pos" & iavser=="neg", "acute",
                               ifelse(iav=="pos" & iavser=="pos", "peak", "late")))) %>% 
  filter(Sample %in% cyto$Sample)
```

## Body Mass Index 
Grecian et al. 2022: estimated body condition by calculating a volume index using the residuals of a linear regression of weight as a function of length × girth^2 (Hammill et al. 1995)  

- Code: 
  - vi_m <- lm(weight_kg ~ length_cm * I(girth_cm^2), dat = dat)
  - summary(vi_m) # adj R squared = 0.487
  - dat$bmi = residuals(vi_m)

Hammill et al. 1995: weight = length X axil1ary girth squared  

```{r}
bmi_model <- lm(weight ~ I(st.length * I(ax.girth^2)), data = metadata)
summary(bmi_model)  # adj R sq = 0.69

bmi <- 
  metadata %>% 
  filter(!weight=="NA" &
         !st.length=="NA" & 
         !ax.girth=="NA") %>% 
  select(Sample) %>% 
  mutate(bmi = bmi_model$residuals)

meta <- 
  metadata %>% 
  left_join(., bmi, by="Sample") %>% 
  mutate(bmi_bin = ifelse(bmi>0, "Robust", "Poor"),
         bmi_bin2 = ifelse(bmi < quantile(bmi, probs=0.25, na.rm=TRUE), "Poor",
                       ifelse(bmi > quantile(bmi, probs=0.25, na.rm=TRUE) & 
                              bmi < quantile(bmi, probs=0.75, na.rm=TRUE),"Average",
                          ifelse(bmi > quantile(bmi, probs=0.75, na.rm=TRUE), "Robust", NA))))

write.csv(x=meta, file="Output Files/metadata_cyto.csv", row.names=FALSE)
```

## Create log, center, scaled data for molt stage, body condition, iav analyses

### Make all non-detects half the minimum concentration detected, or half the limit of detecion (whatever is smallest)
```{r}
minvalue <- 
  cyto_10 %>% 
  pivot_longer(!Sample, names_to="cytokine", values_to="conc") %>% 
  filter(!conc==0) %>% 
  group_by(cytokine) %>% 
  summarize(min=min(conc))

mindc <- 
  read.csv("Input Files/mindc.csv") %>% 
  mutate(cytokine=gsub("[.]", "-", cytokine)) %>% 
  filter(cytokine %in% colnames(cyto_10)) %>% 
  merge(., minvalue) %>% 
  mutate(absmin = pmin(value, min),
         absmin_half = absmin/2)

cyto_scaled <- cyto_10
for (i in 1:nrow(cyto_scaled)) {
  for (j in 1:ncol(cyto_scaled)) {
    {if (cyto_scaled[i,j]==0) {
      k <- which(mindc$cytokine==colnames(cyto_scaled)[j])
      cyto_scaled[i,j] <- mindc$absmin_half[k]
    }
      else{next}
      }
  }
}

cyto_scaled <-
  cyto_scaled %>% 
  mutate_at(2:10, log) %>% 
  mutate_at(2:10, function(x) scale(x, center=TRUE, scale=TRUE))

write.csv(x=cyto_scaled, file="Output Files/cleaned_cytokine_scaled.csv", row.names=FALSE)
```

