---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Exploratory

## Libraries
```{r, warning=FALSE, message=FALSE}
library(mixOmics) # so tidyverse re-writes over select
library(tidyverse)
library(rstatix)
```

## Initial Data Processing 

### Import data & merge
```{r, warning=FALSE, message=FALSE}
cyto <- 
  read.csv("Output Files/cleaned_cytokine_all.csv")
cytobin <-  
  read.csv("Output Files/cleaned_cytokine_bin_all.csv")

cyto_red <-
  read.csv("Output Files/cleaned_cytokine.csv")

meta <- 
  read.csv("Output Files/metadata_cyto.csv") %>% 
  mutate(iav=gsub("neg", "IAV-", iav),
         iav=gsub("pos", "IAV+", iav))

# Merge cytokine & meta data
data <- 
  merge(cyto, meta)

data_red <-
  merge(cyto_red, meta)

databin <- 
  merge(cytobin, meta) 
```

### Ensure variables are correct format
```{r}
data$iav <- as.factor(data$iav)
data$location <- as.factor(data$location)
data$year <- as.factor(data$year)
data$sex <- as.factor(data$sex)
data$molt.stage <- factor(data$molt.stage, levels=c("III", "IV", "V"), ordered=TRUE)
data$bc_bin <- factor(data$bc_bin, levels=c("Poor", "Average", "Robust"))
data$analysis.year <- factor(data$analysis.year)
data$batch <- factor(data$batch, levels=c("1", "2", "3", "4"))
```

## Descriptive Stats 
### Metadata Supplementary Table
```{r}
supplmeta <- 
  data %>% 
  select(Sample, iav, iavser, location, year, sex, molt.stage, 
         bodcond, bc_bin) %>% 
  arrange(., year) %>% 
  mutate(bodcond=paste(round(bodcond, digits=2), " (", bc_bin, ")", sep="")) %>% 
  select(-bc_bin)
write.csv(supplmeta, "Output Files/supplemental_metadata.csv", row.names=FALSE)
```

### Sample sizes across variables
```{r}
table(supplmeta$iav)
table(supplmeta$iavser)
table(supplmeta$location)
table(supplmeta$year)
table(supplmeta$sex)
table(supplmeta$molt.stage)
table(data$bc_bin)
table(data$analysis.year)
```

### Cytokine detections per sample
```{r}
num_samples <- apply(cyto[,2:14], 2, function(x) sum(x>0)) %>%
  data.frame() %>%
  mutate(perc=round((./116)*100, digits=2))
num_samples
```

### Cytokine Ranges
```{r, warning=FALSE}
cyto_pos <- 
  data %>% 
  filter(iav=="IAV+") %>% 
  select(1:14)

std.error <- function(x) sd(x)/sqrt(length(x))

pos_range <- 
  meta %>% 
  select(Sample, iav) %>% 
  filter(iav=="IAV+") %>% 
  merge(., cyto, by="Sample") %>% 
  pivot_longer(!c(iav, Sample), names_to="cytokine", values_to="concentration") %>% 
  filter(!concentration==0) %>% 
  group_by(cytokine) %>% 
  reframe(pos_sampsize=n(),
          mean=mean(concentration),
          se=std.error(concentration),
          median=median(concentration),
          min=min(concentration),
          max=max(concentration)) %>% 
  filter(!duplicated(.)) %>% 
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>% 
  mutate(pos_meanse=paste(mean, "+", se),
         pos_medrange=paste(median, " (", min, " - ", max, ")", sep="")) %>% 
  select(cytokine, pos_sampsize, pos_meanse, pos_medrange)

neg_range <- 
  meta %>% 
  select(Sample, iav) %>% 
  filter(iav=="IAV-") %>% 
  merge(., cyto, by="Sample") %>% 
  pivot_longer(!c(iav, Sample), names_to="cytokine", values_to="concentration") %>% 
  filter(!concentration==0) %>% 
  group_by(cytokine) %>% 
  reframe(neg_sampsize=n(),
          mean=mean(concentration),
          se=std.error(concentration),
          median=median(concentration),
          min=min(concentration),
          max=max(concentration)) %>% 
  filter(!duplicated(.)) %>% 
  mutate(across(where(is.numeric), \(x) round(x, 2))) %>% 
  mutate(neg_meanse=paste(mean, "+", se),
         neg_medrange=paste(median, " (", min, " - ", max, ")", sep="")) %>% 
  select(cytokine, neg_sampsize, neg_meanse, neg_medrange)

table <- merge(neg_range, pos_range, by="cytokine", all.x=TRUE)
write.csv(table, "Output Files/manuscript_cytokinetable.csv", row.names=FALSE)
```

## PCA
```{r}
pca <- 
  prcomp(data_red[,2:10], center=TRUE, scale=TRUE)

pca_df <- 
  data.frame(
    x=pca$x[,1],
    y=pca$x[,2],
    molt.stage=data$molt.stage,
    year=factor(data$year),
    location=data$location,
    sex=data$sex,
    iav=data$iav,
    iavser=data$iavser,
    bc_bin=data$bc_bin,
    analysis.year=data$analysis.year,
    batch=data$batch
    )

pca_labelx <-
  paste("PC1 (", 
        round(abs(summary(pca)$importance[2,1]*100), digits=0),
        "%)",
        sep="")
pca_labely <-
  paste("PC2 (", 
        round(abs(summary(pca)$importance[2,2]*100), digits=0),
        "%)",
        sep="")
```

### Molt Stage
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=molt.stage)) +
  geom_point() +
  labs(x=pca_labelx,
       y = pca_labely) +
  scale_color_manual("Molt Stage",
                     values=c("#9d9d9d","#008ba2", "#000000")) +
  theme_bw() +
  theme(panel.grid=element_blank())
```

### Year
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=year)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())

table(data$year, data$iav)
```

#### What's up with year?
Similar # detections per sample as in other years (det.samp), and similar total cytokine concentrations per sample as in other years (sum.samp). 
```{r}
yeartab <-
  data.frame(table(data$year))
  
data %>% 
  select(Sample, 2:10, year) %>% 
  pivot_longer(!c(Sample, year), names_to="cytokine", values_to="concentration") %>%
  filter(!concentration==0) %>% 
  group_by(year) %>% 
  summarize(sum = sum(concentration),
            mean = mean(concentration),
            median = median(concentration),
            n.det = n(),
            n.cyto = length(unique(cytokine))) %>% 
  data.frame() %>% 
  merge(., yeartab, by.x="year", by.y="Var1") %>% 
  mutate(det.samp = n.det/Freq,
         sum.samp = sum/Freq)
```

### Analysis Year
Year that samples were analyzed, "batches"
```{r}
analysisyear_plot <-
  ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=analysis.year)) +
  geom_point() +
  stat_ellipse() +
  scale_color_manual(values=c("black", "slategray3", "grey60"), 
                     name = "Analysis Year") +
  labs(x = pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = "bottom")
analysisyear_plot

ggsave(filename="Figures/analysisyear_pca.jpeg", analysisyear_plot, 
       height=6, width=6, units="in")

biplot(pca)
```

#### What's up with analysis year?
Similar # detections per sample as in other years (det.samp), and similar total cytokine concentrations per sample as in other years (sum.samp). 
```{r}
analysisyeartab <-
  data.frame(table(data$analysis.year))
  
batchsummary <-
  data %>% 
  select(Sample, 2:14, analysis.year) %>% 
  pivot_longer(!c(Sample, analysis.year), names_to="cytokine", values_to="concentration") %>%
  filter(!concentration==0) %>% 
  group_by(analysis.year, cytokine) %>% 
  summarize(sum = sum(concentration),
            mean = mean(concentration),
            median = median(concentration),
            n.det = n(),
            n.cyto = length(unique(cytokine)),
            max = max(concentration)) %>% 
  data.frame() %>% 
  merge(., analysisyeartab, by.x="analysis.year", by.y="Var1") %>% 
  mutate(det.samp = n.det/Freq,
         sum.samp = sum/Freq)

# Mean cytokine concentration
ggplot(batchsummary, aes(x=cytokine, y=mean, fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Mean Concentration") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))

ggplot(batchsummary, aes(x=cytokine, y=log(mean), fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Log Tranformed Mean Concentration") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))


# Median cytokine concentration
ggplot(batchsummary, aes(x=cytokine, y=median, fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Median Concentration") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))

ggplot(batchsummary, aes(x=cytokine, y=log(median), fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Log Transformed Median Concentration") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))


# Sum of cytokine concentrations, sum/sample
ggplot(batchsummary, aes(x=cytokine, y=log(sum), fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Log Transformed Sum Concentration") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))
 
ggplot(batchsummary, aes(x=cytokine, y=log(sum.samp), fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Log Transformed (Sum Concentration/Samples Analyzed)") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))


# Detections and proportion of samples with detections
ggplot(batchsummary, aes(x=cytokine, y=n.det, fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Number of detections") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))

ggplot(batchsummary, aes(x=cytokine, y=det.samp, fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Number of detections per Sample") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))


# Max detection of each cytokine
ggplot(batchsummary, aes(x=cytokine, y=max, fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Max Detection") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))

ggplot(batchsummary, aes(x=cytokine, y=log(max), fill=analysis.year)) +
  geom_col(position="dodge") +
  labs(y = "Log Transformed Max Detection") +
  scale_fill_manual(values=c("slategray", "slategray3", "slategray2"))
```


### Batch
Batches that samples were analyzed in
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=batch)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
```

### Location
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=location)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
```

### Sex
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=sex)) +
  geom_point() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
```

### Body Condition
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=bc_bin)) +
  geom_point() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
```

### IAV
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=iav)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
```

### IAV Serology
```{r}
ggplot(pca_df, 
       aes(x=x, 
           y=y,
           col=iavser)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
```

## Is proportion of IAV+/IAV- similar across variables?

### Location
```{r}
tryCatch({
  chisq.test(table(data$location, data$iav))
}, warning = function(w) {
  fisher.test(table(data$location, data$iav))
})
```

### Molt Stage
```{r}
tryCatch({
  chisq.test(table(data$molt.stage, data$iav))
}, warning = function(w) {
  fisher.test(table(data$molt.stage, data$iav))
})

data.frame(table(data$iav, data$molt.stage)) %>% 
  pivot_wider(names_from="Var1", values_from="Freq") %>% 
  mutate(proppos = (`IAV+`/(`IAV+` + `IAV-`))*100)


# Pairwise post-hoc
pairwise_fisher_test(table(data$molt.stage, data$iav), p.adjust.method="holm")
```

### Body Condition
```{r}
tryCatch({
  chisq.test(table(data$bc_bin, data$iav))
}, warning = function(w) {
  fisher.test(table(data$bc_bin, data$iav))
})
```

### Analysis Year
```{r}
tryCatch({
  chisq.test(table(data$analysis.year, data$iav))
}, warning = function(w) {
  fisher.test(table(data$analysis.year, data$iav))
})

# Pairwise post-hoc
pairwise_fisher_test(table(data$analysis.year, data$iav), p.adjust.method="holm")
```

