---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Molt stage ~ cytokines

## Libraries
```{r, warning=FALSE, message=FALSE}
library(vegan)
library(mixOmics)
library(tidyverse)
library(ggfortify)
library(cramer)
library(mlr3)
library(paradox)
library(mlr3tuning)
library(rpart)
library(rpart.plot)
library(pairwiseAdonis)
library(ggmosaic)
library(EnvStats)
library(gridExtra)
library(grid)
```

## Data
```{r, message=FALSE, warning=FALSE}
cyto <- read.csv("Output Files/cleaned_cytokine.csv")
cytobin <- read.csv("Output Files/cleaned_cytokine_bin.csv")
cytoscale <- read.csv("Output Files/cleaned_cytokine_scaled.csv")
meta <- read.csv("Output Files/metadata_cyto.csv")

# Merge cytokine & metadata 
data <- merge(cyto, meta, by="Sample") %>% 
  filter(!is.na(molt.stage))

databin <- 
  merge(cytobin, meta, by="Sample") %>% 
  filter(!is.na(molt.stage))

datascale <- 
  merge(cytoscale, meta, by="Sample") %>% 
  filter(!is.na(molt.stage))
```


## PCA
```{r}
PCA_scaled_molt <- prcomp(datascale[,2:10], center=FALSE, scale=FALSE)
```

### Visualize
```{r}
PCA_scaled_molt_df <- data.frame(x=PCA_scaled_molt$x[,1],
                                 y=PCA_scaled_molt$x[,2],
                                 molt.stage=datascale$molt.stage)

pca_labelx <-
  paste("PC1 (", 
                round(abs(summary(PCA_scaled_molt)$importance[2,1]*100), digits=2),
                "%)",
                sep="")
pca_labely <-
  paste("PC2 (", 
                round(abs(summary(PCA_scaled_molt)$importance[2,2]*100), digits=2),
                "%)",
                sep="")

PCA_scaled_molt_ggplot <- 
  ggplot(PCA_scaled_molt_df, 
         aes(x=x, 
             y=y,
             col=molt.stage)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely,
       name="Molt Stage") +
  scale_color_manual("Molt Stage",
                     values=c("#9d9d9d","#008ba2", "#000000")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position.inside=c(0.1, 0.1))

PCA_scaled_molt_ggplot
```

## PERMANOVA

### Cytokine Presence/Absence
From McCosker et al. 2024:  

- A dissimilarity matrix was then created using betadiver from package vegan (v. 2.5-4; Oksanen et al.
2018) with dissimilarity deﬁned as 1 – Sørensen’s index (Sørensen 1948).

#### Create similarity matrix (Sorensen)
```{r}
databin_perm <- 
  databin %>% 
  select(2:10)

databin_dist <- 
  betadiver(databin_perm, method=11)

plot(databin_dist)

# Dissimilarity summary stats
mean(1-databin_dist); min(1-databin_dist); max(1-databin_dist); median(1-databin_dist)
```

#### PERMANOVA with dissimilarity (1-Sorensen)
R2 = partial R2 = factor (molt stage) is explaining at least R2 * 100% of variation in variables (cytokines)
```{r}
databin_adonis <- 
  adonis2((1-databin_dist) ~ molt.stage, data=databin, permutations=4999)
databin_adonis

# R2
databin_adonis$R2[1]*100
```

#### Visualizing multivariate groups
Spider plot illustrating group dissimilarity and group centroid locations (molt stage) using presence/absence data to compare cytokine detections across molt stages.

##### Principal coordinates analysis 
```{r, warning=FALSE}
# Use 1 - Sorensen's for dissimilarity 
databin_pco <- cmdscale(1-databin_dist, eig=TRUE)
```

##### Plot
```{r}
plot(databin_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)

ordispider(jitter(databin_pco$points, factor=1), factor(databin$molt.stage), show.groups="III", 
           col="deepskyblue", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(databin_pco$points, factor=1), factor(databin$molt.stage), show.groups="IV", 
           col="firebrick", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(databin_pco$points, factor=1), factor(databin$molt.stage), show.groups="V", 
           col="darkmagenta", label = TRUE, display="sites", lwd=1.5, cex=1.2)
```

#### Homogeneity of group dispersions
```{r}
disper <- 
  betadisper(databin_dist, group=databin$molt.stage, type="centroid")
anova(disper)
boxplot(disper)
```

#### Posthoc pairwise tests
```{r}
pairwise.adonis(databin_dist, databin$molt.stage)
```


### Cytokine Concentration
From McCosker et al. 2024:  

- The adonis function was used to conduct a permutational multivariate analysis of variance (PERMANOVA) on group (method) dissimilarity. Group dispersions weremodeled with betadisper to determine whether the assumption of equal dispersions had been met under adonis and principal coordinate analysis was used relative to PERMANOVA results to visualize group dispersions around the centroid.  
- Bray-Curtis dissimilarity used  
  - well-suited to species abundance because it ignores variables that have zeros for both objects (joint absences)
```{r}
data_perm <-
  data %>% 
  select(2:10)

data_dist <- 
  vegdist(data_perm, method="bray")

data_adonis <- adonis2(data_dist ~ molt.stage, data=data, 
                       permuations=4999)
data_adonis

# R2
data_adonis$R2[1]*100
```

#### Visualizing multivariate groups

##### Principal coordinates analysis 
```{r, warning=FALSE}
# Use Bray-Curtis dissimilarity matrix
data_pco <- cmdscale(data_dist, eig=TRUE)
```

##### Plot
```{r}
plot(data_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)

ordispider(jitter(data_pco$points, factor=1), factor(data$molt.stage), show.groups="III", 
           col="deepskyblue", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(data_pco$points, factor=1), factor(data$molt.stage), show.groups="IV", 
           col="firebrick", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(data_pco$points, factor=1), factor(data$molt.stage), show.groups="V", 
           col="darkmagenta", label = TRUE, display="sites", lwd=1.5, cex=1.2)
```

#### Homogeneity of group dispersions
```{r}
disper <- 
  betadisper(data_dist, group=data$molt.stage, type="centroid")
anova(disper)
boxplot(disper)
```

#### Posthoc pairwise tests
```{r}
pairwise.adonis(data_dist, data$molt.stage)
```


## Cramer Test 
Multivariate composition of the cytokine pattern between each molt stage pairwise comparison It can handle multiple variables, potentially complex interactions and correlations.  
```{r}
# Separate data by molt stage
three <- 
  datascale %>% 
  filter(molt.stage=="III") %>% 
  select(2:10) %>% 
  data.matrix()

four <- 
  datascale %>% 
  filter(molt.stage=="IV") %>% 
  select(2:10) %>% 
  data.matrix()

five <- 
  datascale %>% 
  filter(molt.stage=="V") %>% 
  select(2:10) %>% 
  data.matrix()

# Run Cramer Tests!
cramer.test(three, four)
cramer.test(four, five)
cramer.test(three, five) ###
```

## Cytokine Importance
The default splitting method for classification is gini index. Optional parameters:  

- minbucket: smallest # observations allowed in terminal node. Small terminal nodes likely to be biased by single outlier.  
- minsplit: smallest # observations in the parent node that can be further split. Default=20.  
- maxdepth: prevents tree from growing past certain depth/height. Default=30.  
- cp: complexity pararmeter: minimum improvement in the model needed at each node. Based on the cost complexity of the model (measures mis-classification rate of potential branches). Small cp will allow for more complex models.  
  - changing cp from default to 0.001 does not change tree*  

Cross Validation  

- split training set in 2+ parttitions & creating model for each partition. Partition acts as validation set, and remaining set is training data. Common form is k-fold:  
  - The models are always trained on k - 1 of the folds, with the remaining fold being used as test data; this process is repeated until each fold has acted exactly once as test set.
  
Classification Ability  

- ROC AUC (range 0-1)  
  - 0.5 = random guessing  
  - 1 = indicates perfect performance  
  - A score slightly above 0.5 shows that a model has at least "some" (albeit small) predictive power. This is generally inadequate for any real applications.  

- Classification Error  

  - compares observed to predicted classifications  
  - range from 0-1, with smaller numbers = less error, better model  

- Resources:  
  - https://mlr3book.mlr-org.com/chapters/chapter4/hyperparameter_optimization.html#sec-tuning-instance  
  - https://mlr3book.mlr-org.com/chapters/chapter3/evaluation_and_benchmarking.html#sec-resampling  

```{r, message=FALSE}
# Prepare data for CART
data_cart <- 
  datascale %>% 
  select(2:10, molt.stage) %>% 
  mutate(molt.stage=factor(molt.stage, levels=c("III", "IV", "V")))
```


### Cross Validation
```{r}
# Creating task and learner
task <- as_task_classif(molt.stage ~ ., 
                        data = data_cart)
        
task <- task$set_col_roles(cols="molt.stage",
                           add_to="stratum")

learner <- lrn("classif.rpart",
               predict_type = "prob",
               maxdepth = to_tune(2, 5),
               minbucket = to_tune(1, 7),
               minsplit = to_tune(1, 30)
               )

# Define tuning instance - info ~ tuning process
instance <- ti(task = task,
               learner = learner,
               resampling = rsmp("cv", folds = 10),
               measures = msr("classif.ce"),
               terminator = trm("none")
               )

# Define how to tune the model
tuner <- tnr("grid_search", 
             batch_size = 10
             )

# Trigger the tuning process
#tuner$optimize(instance)
 

# optimal values: 
  # maxdepth = 2
  # minbucket = 6
  # minsplit = 30
  # classif.ce = 0.3170455
```

### Run model with optimized parameteres
Leaves out molt stage III - too few samples for CART alogrithm
```{r}
cart_model <- rpart(formula=molt.stage ~ .,
                    data=data_cart,
                    method="class",
                    maxdepth=2,
                    minbucket=6,
                    minsplit=30)


# plot optimized tree
rpart.plot(cart_model)
cart_model$variable.importance

# IL-8 most important cytokine
```

### Plot important cytokines
#### Variable Importance
```{r}
varimp <- 
  data.frame(imp=cart_model$variable.importance) %>% 
  rownames_to_column() %>% 
  rename("variable" = rowname) %>% 
  arrange(imp) %>%
  mutate(variable=gsub("\\.","-",variable),
         variable = forcats::fct_inorder(variable))

varimp_plot <- 
  ggplot(varimp) + 
  geom_segment(aes(x = variable, 
                   y = 0, 
                   xend = variable, 
                   yend = imp), 
               linewidth = 0.5, 
               alpha = 0.7) +
  geom_point(aes(x = variable, 
                 y = imp), 
             size = 1, 
             show.legend = F,
             col="black") +
  labs(x="Cytokine", 
       y="Variable Importance") +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=10, color="black"))
```

#### Proportions
```{r, warning=FALSE}
prop.table(table(databin$molt.stage, databin$IL.8), margin=1)*100

il8_prop <- 
  databin %>% 
  mutate(IL.8=gsub("1", "P", IL.8),
         IL.8=gsub("0", "A", IL.8),
         IL.8=factor(IL.8, levels=c("P", "A"))) %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.8,molt.stage), fill=IL.8)) +
  scale_fill_manual(values=c("P"="gray60", "A"="lightgray")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position="none",
        text = element_text(size=10),
        plot.title = element_text(hjust = 0.5))

il8_prop
```

#### Concentration
```{r}
il8 <-
  data %>% 
  select(molt.stage, IL.8) %>% 
  filter(!IL.8==0) %>% 
  ggplot(., aes(x=molt.stage, y=log(IL.8))) +
  geom_boxplot(color="black", fill="gray60", lwd=0.25,
               outliers=FALSE) +
  geom_point(size=0.75, position=position_jitter(width=0.2)) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=3) +
  theme_classic() +
  theme(text = element_text(size=10),
         panel.border=element_rect(color="black", fill=NA, linewidth=0.5))
il8

data %>% 
  select(molt.stage, IL.8) %>% 
  filter(!IL.8==0) %>% 
  group_by(molt.stage) %>% 
  summarize(median=median(IL.8))
```

#### Combine
```{r}
bottom <- grid.arrange(il8_prop, il8, ncol=2,
                       left=textGrob("IL-8", rot=90, gp=gpar(fontsize=10)))

jpeg("Figures/cytoimportance_molt.jpeg", width=7, height=6, units="in", res=300)
all <- grid.arrange(varimp_plot, bottom, 
                    ncol=1, heights=c(1,1.5),
                    bottom=textGrob("Molt Stage", gp=gpar(fontsize=10)))
grid.polygon(x=c(0, 0, 1, 1),
             y=c(0, 0.62, 0.62, 0),
             gp=gpar(fill=NA))
grid.polygon(x=c(0.02, 0.02, 0.97, 0.97),
             y=c(0.92, 0.97, 0.97, 0.92),
             gp=gpar(fill=NA))
dev.off()
```


## How different groups are
Methods: 

- Partial least squares regression (PLSR) figures  
- 2-dimensional Kolmogorov-Smirnov test  
  - not sure this one makes sense!

### PLSDA 
PLSDA is a supervised dimensionality reduction method that is similar to PCA but incorporates class labels to maximize the separation between groups.  
One advantage of PLS is that it has fewer assumptions and hence it can use predictor variables that are collinear and not independent.  
The aim of PLS-DA is to discriminate the sample groups, rather than maximizing the variance. When the maximization of the variance corresponds to a discrimination of the classes, then this is awesome, but otherwise, it means that the major source of variation is probably not mainly due to the sample group differences.

#### Prepare data
```{r}
pls_x <-
 data_cart[,1:9] 

pls_y <- 
  data_cart %>% 
  select(molt.stage)
```

#### PLSDA Model
```{r}
plsda_model <- mixOmics::plsda(pls_x,
                               pls_y$molt.stage,
                               scale=TRUE,
                               ncomp=9)
```

#### Visualizations
```{r}
# Set x and y axis labels based on model loadings
labelx <- paste("X-variate 1: ", 
                round(abs(plsda_model$prop_expl_var$X[1]*100), digits=0),
                "% expl. var",
                sep="")

labely <- paste("X-variate 2: ", 
                round(abs(plsda_model$prop_expl_var$X[2]*100), digits=0),
                "% expl. var",
                sep="")

# ggplot2
plsda_ggplot <- 
  data.frame(x = plsda_model$variates$X[,1], 
             y = plsda_model$variates$X[,2],
             molt.stage = plsda_model$Y)

molt_plsda_ggplot <- 
  ggplot(plsda_ggplot, 
         aes(x=x, 
             y=y,
             col=molt.stage)) +
  geom_point() +
  stat_ellipse() +
  labs(x=labelx,
       y = labely,
       name="Molt Stage") +
  scale_color_manual("Molt Stage",
                     values=c("#9d9d9d","#008ba2","black")) +
  theme_bw() +
  theme(panel.grid=element_blank())
molt_plsda_ggplot
```

#### Cytokine Contributions
```{r}
plsda_loadings <- 
  plsda_model$loadings$X %>% 
  data.frame() %>% 
  select(1:2) %>% 
  arrange(comp1)
plsda_loadings

biplot(plsda_model, ind.names=FALSE, legend.title="Molt Stage")

plotVar(plsda_model)
```

