---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Body condition ~ cytokines

## Libraries
```{r, warning=FALSE, message=FALSE}
library(vegan)
library(mixOmics)
library(tidyverse)
library(pairwiseAdonis)
library(mlr3)
library(mlr3tuning)
library(rpart)
library(rpart.plot)
library(ggmosaic)
library(ggpubr)
library(EnvStats)
library(gridExtra)
library(grid)
library(PLSDAbatch)
```

## Data
```{r, message=FALSE, warning=FALSE}
cyto <- 
  read.csv("Output Files/cleaned_cytokine.csv")
cytobin <- 
  read.csv("Output Files/cleaned_cytokine_bin.csv")

meta <- 
  read.csv("Output Files/metadata_cyto.csv") %>% 
  mutate(iav=gsub("neg", "IAV-", iav),
         iav=gsub("pos", "IAV+", iav))

# Merge cytokine & metadata 
data <- 
  merge(cyto, meta, by="Sample") %>% 
  filter(!is.na(bc_bin)) %>% 
  mutate(bc_bin = factor(bc_bin, levels=c("Poor", "Average", "Robust")))

databin <- 
  merge(cytobin, meta, by="Sample") %>% 
  filter(!is.na(bc_bin)) %>% 
  mutate(bc_bin = factor(bc_bin, levels=c("Poor", "Average", "Robust")))
```


## Body Condition Range & Bins
```{r}
quant <- quantile(meta$bodcond, na.rm=TRUE)

ggplot(meta, aes(x=bodcond)) +
  geom_histogram(bins=10, color="black", fill="gray80") +
  labs(x="Body Condition Index", y="Frequency") +
  geom_vline(xintercept=c(quant[2], quant[3], quant[4]), colour="red", linewidth=1) +
  theme_classic()
```


## PERMANOVA

### Cytokine Presence/Absence

#### Create similarity matrix (Sorensen)
Dissimilarity = 1-Sorensen
```{r}
databin_dist <- 
  databin %>% 
  select(2:10) %>% 
  betadiver(., method=11)

plot((1-databin_dist))

# Dissimilarity summary stats
mean(1-databin_dist); min(1-databin_dist); max(1-databin_dist); median(1-databin_dist)
```

#### Run PERMANOVA
Use restricted permutation test - data are not exchangeable between analysis years so permute within groups defined by strata.  
** Not sig
```{r}
# define permutations & strata
permute_databin <- 
  how(plots=Plots(strata=databin$analysis.year, type="none"), nperm=4999)

# run permanova
adonis2((1-databin_dist) ~ bc_bin, 
        data=databin,
        permutations=permute_databin)
```

#### Homogeneity of group dispersions
** Robust vs Poor & Robust vs Average
```{r}
disper <- 
  betadisper((1-databin_dist), group=databin$bc_bin, type="centroid")

anova(disper)
TukeyHSD(disper)
boxplot(disper)
```

#### Visualization
Principal Coordinates Analysis
```{r, warning=FALSE}
# Use 1 - Sorensen's for dissimilarity 
databin_pco <- cmdscale((1-databin_dist), eig=TRUE)

plot(databin_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)
ordiellipse(ord = databin_pco, 
            groups = factor(data$bc_bin), 
            display = "sites", 
            col = c("deepskyblue", "grey", "darkseagreen"),
            lwd = 2,
            label = TRUE)

plot(databin_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)
ordispider(ord = databin_pco,
           groups = factor(data$bc_bin),
           display = "sites",
           col = c("deepskyblue", "grey", "darkseagreen"),
           lwd=2,
           label=TRUE)
```


### Cytokine Concentration
#### Dissimilarity matrix
```{r}
data_dist <-
  data %>% 
  select(2:10) %>% 
  vegdist(., method="bray")
```

#### Run PERMANOVA
** Not sig
```{r}
# define permutations & strata
permute_data <- how(plots=Plots(strata=data$analysis.year, type="none"), nperm=4999)

# run permanova
adonis2(data_dist ~ bc_bin, 
        data=data,
        permutations=permute_data)
```

#### Homogeneity of group dispersions
** Not sig
```{r}
disper <- 
  betadisper(data_dist, group=data$bc_bin, type="centroid")

anova(disper)
boxplot(disper)
```

#### Visualization
```{r, warning=FALSE}
# Use Bray-Curtis dissimilarity matrix
data_pco <- cmdscale(data_dist, eig=TRUE)

plot(data_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)
ordiellipse(ord = data_pco, 
            groups = factor(data$bc_bin), 
            display = "sites", 
            col = c("deepskyblue", "grey", "darkseagreen"),
            lwd = 2,
            label = TRUE)

plot(data_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)
ordispider(ord = data_pco,
           groups = factor(data$bc_bin),
           display = "sites",
           col = c("deepskyblue", "grey", "darkseagreen"),
           lwd=2,
           label=TRUE)
```


## Cytokine Importance
### Prepare data
```{r, message=FALSE}
data_cart <- 
  data %>% 
  select(2:10, bc_bin) %>% 
  mutate(bc_bin=factor(bc_bin, levels=c("Poor", "Average", "Robust")))
```

### Cross Validation
See reference: https://mlr3book.mlr-org.com/
```{r}
# Creating task and learner
task <- as_task_classif(bc_bin ~ ., 
                        data = data_cart)

task <- task$set_col_roles(cols="bc_bin",
                           add_to="stratum")

# Minimum body condition group size = 25
learner <- lrn("classif.rpart",
               predict_type = "prob",
               maxdepth = to_tune(2, 5),
               minbucket = to_tune(1, 25),
               minsplit = to_tune(1, 30)
               )

# Define tuning instance - info ~ tuning process
instance <- ti(task = task,
               learner = learner,
               resampling = rsmp("cv", folds = 10),
               measures = msr("classif.ce"),
               terminator = trm("none")
               )

# Define how to tune the model
tuner <- tnr("grid_search", 
             batch_size = 10
             )

# Trigger the tuning process
#tuner$optimize(instance)
 
# optimal values: 
  # maxdepth = 4
  # minbucket = 3
  # minsplit = 20
  # classif.ce = 0.5058159
```

### Run CART 
```{r}
cart_model <- rpart(formula=bc_bin ~ .,
                    data=data_cart,
                    method="class",
                    maxdepth=4,
                    minbucket=3,
                    minsplit=20)

# plot optimized tree
rpart.plot(cart_model)
cart_model$variable.importance 
  # IFNg is most important cytokine
```

### Plot 
#### Variable Importance
```{r}
varimp <- 
  data.frame(imp=cart_model$variable.importance) %>% 
  rownames_to_column() %>% 
  rename("variable" = rowname) %>% 
  arrange(imp) %>%
  mutate(variable=gsub("\\.","-",variable),
         variable = forcats::fct_inorder(variable))

varimp_plot <- 
  ggplot(varimp) + 
  geom_segment(aes(x = variable, 
                   y = 0, 
                   xend = variable, 
                   yend = imp), 
               linewidth = 0.5, 
               alpha = 0.7) +
  geom_point(aes(x = variable, 
                 y = imp), 
             size = 1, 
             show.legend = F,
             col="black") +
  labs(x="Cytokine", 
       y="Variable Importance") +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=10, color="black"))
varimp_plot
```

#### Proportions
```{r}
ifng_prop <- 
  databin %>% 
  mutate(IFNg=gsub("1", "Present", IFNg),
         IFNg=gsub("0", "Absent", IFNg),
         IFNg=factor(IFNg, levels=c("Present", "Absent"))) %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IFNg, bc_bin), fill=IFNg)) +
  scale_fill_manual(values=c("Present"="gray60", "Absent"="lightgray"),
                    name="Cytokine") +
  scale_y_continuous(labels = scales::percent) +
  labs(y="Proportion of Samples") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        legend.position="bottom",
        text = element_text(size=10),
        plot.title = element_text(hjust = 0.5))

cyto_leg <- 
  get_legend(ifng_prop)

ifng_prop <-
  ifng_prop +
  theme(legend.position="none")
```

#### Concentration
```{r}
ifng <-
  data %>% 
  select(bc_bin, IFNg) %>% 
  filter(!IFNg==0) %>% 
  ggplot(., aes(x=bc_bin, y=log(IFNg))) +
  geom_boxplot(color="black", fill="gray60", lwd=0.25,
               outliers=FALSE) +
  geom_point(size=0.75, position=position_jitter(width=0.2)) +
  labs(x=NULL, y="log(Concentration pg/mL)") +
  stat_n_text(size=3) +
  theme_classic() +
  theme(text = element_text(size=10),
         panel.border=element_rect(color="black", fill=NA, linewidth=0.5))
```

#### Combine
```{r}
cyto <- grid.arrange(ifng_prop,ifng,
                    ncol=2, 
                    top = textGrob(label="IFNg"))

bottom <- grid.arrange(cyto, cyto_leg, ncol=1, 
                     heights=c(3,0.5))

jpeg("Figures/cytoimportance_bodycondition.jpeg", width=8, height=6, units="in", res=300)
all <- grid.arrange(varimp_plot, bottom, 
                    heights=c(1.25,2),
                    bottom=textGrob("Body Condition", gp=gpar(fontsize=10)))
grid.polygon(x=c(0, 0, 1, 1),
             y=c(0, 0.68, 0.68, 0),
             gp=gpar(fill=NA))
grid.polygon(x=c(0.02, 0.02, 0.97, 0.97),
             y=c(0.955, 0.99, 0.99, 0.955),
             gp=gpar(fill=NA))
dev.off()
```


## PLS-DA

### Data 
```{r}
pls_x <-
  data[,2:10]

pls_y <- 
  data %>% 
  select(bc_bin, analysis.year) %>% 
  mutate_all(as.factor)
```

### Transform data 
CLR = Centered log-ratio transformation
```{r}
pls_x_clr <- 
  logratio.transfo(X = pls_x, logratio = 'CLR', offset = 0.01)

class(pls_x_clr) = 'matrix'
```

### Detect batch effects
Reference: https://evayiwenwang.github.io/PLSDAbatch_workflow/ 
```{r}
pca_before <- 
  mixOmics::pca(pls_x_clr, scale=TRUE)

plotIndiv(pca_before, 
          group=data$analysis.year,
          pch=20,
          legend = TRUE, legend.title = "Analysis Year",
          title = "Before Batch Correction",
          ellipse = TRUE, ellipse.level=0.95)

Scatter_Density(object = pca_before,
                batch = pls_y$analysis.year,
                trt = pls_y$bc_bin,
                trt.legend.title = "Body Condition",
                color.set = c("#9d9d9d","#008ba2","black"))
```

#### Estimate # variable components
Use PLSDA from mixOmics with only treatment to choose number of treatment components to preserve.  
Number = that explains 100% of variance in outcome matrix Y.
```{r}
trt_comp <- 
  plsda(X = pls_x_clr,
        Y = pls_y$bc_bin, 
        ncomp=9)

trt_comp$prop_expl_var
sum(trt_comp$prop_expl_var$Y[seq_len(3)]) # preserve 3 comp
```

#### Estimate # batch components
Use PLSDA_batch with both treatment and batch to estimate optimal number of batch components to remove.  
For ncomp.trt, use # components found above. Samples are not balanced across batches x treatment, using balance = FALSE.  
Choose # that explains 100% variance in outcome matrix Y. 
```{r}
batch_comp <- 
  PLSDA_batch(X = pls_x_clr,
              Y = pls_y$bc_bin,
              Y.bat = pls_y$analysis.year,
              balance = FALSE,
              ncomp.trt = 3, ncomp.bat = 9)

batch_comp$explained_variance.bat
sum(batch_comp$explained_variance.bat$Y[seq_len(3)]) # preserve 3 comp
```

#### Correct for batch effects
Use optimal number of components determined above.  
Treatment = 3, Batch = 3
```{r}
plsda_batch <-
  PLSDA_batch(X = pls_x_clr,
              Y = pls_y$bc_bin,
              Y.bat = pls_y$analysis.year,
              balance = FALSE,
              ncomp.trt = 3, ncomp.bat = 3)

data_batch <-
  plsda_batch$X.nobatch
```

#### Assessing batch correction
```{r}
after_batch <- 
  mixOmics::pca(data_batch, scale=TRUE)

plotIndiv(after_batch, 
          group=pls_y$analysis.year,
          pch=20, 
          legend = TRUE, legend.title = "Analysis Year",
          title = "After Batch Correction",
          ellipse = TRUE, ellipse.level = 0.95)
```

### PLSDA on Batch-Corrected Data
```{r}
plsda_model <- 
  plsda(data_batch,
        pls_y$bc_bin,
        scale=TRUE)

plotIndiv(plsda_model,
          group=pls_y$bc_bin,
          pch=20, 
          legend = TRUE, legend.title = "Body Condition",
          ellipse = TRUE, ellipse.level = 0.95)
```

### Visualization
#### Set x and y axis labels 
```{r}
labelx <- 
  paste("X-variate 1: ", 
        round(abs(plsda_model$prop_expl_var$X[1]*100), digits=0),
        "% expl. var",
        sep="")

labely <- 
  paste("X-variate 2: ", 
        round(abs(plsda_model$prop_expl_var$X[2]*100), digits=0),
        "% expl. var",
        sep="")
```

#### Create plot data frame
```{r}
plsda_ggplot <- 
  data.frame(x = plsda_model$variates$X[,1], 
             y = plsda_model$variates$X[,2],
             bc_bin = plsda_model$Y)
```

#### Plot
```{r}
bc_plsda_ggplot <- 
  ggplot(plsda_ggplot, 
         aes(x=x, 
             y=y,
             col=bc_bin)) +
  geom_point() +
  stat_ellipse() +
  labs(x=labelx,
       y = labely,
       name="Body Condition") +
  scale_color_manual("Body Condition",
                     values=c("#9d9d9d","#008ba2", "black")) +
  theme_bw() +
  theme(panel.grid = element_blank())

bc_plsda_ggplot
```

### Cytokine Contributions
```{r}
plsda_loadings <- 
  plsda_model$loadings$X %>% 
  data.frame() %>% 
  select(1:2) %>% 
  arrange(comp1)
plsda_loadings

biplot(plsda_model, ind.names=FALSE, legend.title="Body Condition")

plotVar(plsda_model)
```