---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Body condition ~ cytokines

## Load required libraries
```{r, warning=FALSE, message=FALSE}
library(mixOmics)
library(tidyverse)
library(ggfortify)
library(cramer)
library(mlr3)
library(paradox)
library(mlr3tuning)
library(rpart)
library(rpart.plot)
```

## Read in data 
```{r, message=FALSE, warning=FALSE}
cyto <- read.csv("Output Files/cleaned_cytokine.csv")
meta <- read.csv("Output Files/metadata_cyto.csv")
mindc <- read.csv("Input Files/mindc.csv") %>% 
  mutate(half=value/2)

# Merge cytokine & metadata 
data <- merge(cyto, meta, by="Sample")

# Tidy data, remove cytokines with < 10% (11) detections, remove pups with NA bc
data_tidy <-
  data %>%
  select(2:14) %>% 
  select_if(function(x) sum(x>0) > 11) %>% 
  mutate(bc=data$bc_bin,
         bmi=data$bmi_bin,
         bmi2=data$bmi_bin2,
         sample=data$Sample)
```

## Log-transforme scaled,center cytokine data
```{r}
# Make all non-detects half the limit of detection
data_half <- data_tidy
for (i in 1:nrow(data_half)) {
  for (j in 1:9) {
     {if (data_half[i,j]==0) {
       k <- which(mindc==colnames(data_half)[j])
       data_half[i,j] <- mindc[k,3]
     }
       else{next}
       }
  }
}

# Log transform, scale, and center data
data_log <- 
  data_half %>% 
  mutate_at(1:9, log)

data_scaled <- 
  data_half %>% 
  mutate_at(1:9, log) %>% 
  mutate_at(1:9, function(x) scale(x, center=TRUE, scale=TRUE))

data_scaled_bc <-
  data_scaled %>% 
  filter(!bc=="NA")

data_scaled_bmi <-
  data_scaled %>% 
  filter(!bmi=="NA")
```

## PCAs
```{r, warning=FALSE}
PCA_scaled_bc <- 
  prcomp(~IFNg+IL.2+IL.6+IL.7+IL.8+IL.10+IL.15+IL.18+KC.like, 
         data=data_scaled_bc,
         center=FALSE, 
         scale=FALSE)
autoplot(PCA_scaled_bc, data=data_scaled_bc, shape="bc") +
         scale_shape_manual(values=c(1,19,15), name="Body Condition") +
         theme_bw() +
         theme(panel.grid=element_blank(), text=element_text(size=8),
               legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
               legend.position="bottom")

PCA_scaled_bmi <-
  prcomp(~IFNg+IL.2+IL.6+IL.7+IL.8+IL.10+IL.15+IL.18+KC.like, 
         data=data_scaled_bmi,
         center=FALSE, 
         scale=FALSE)
autoplot(PCA_scaled_bmi, data=data_scaled_bmi, shape="bmi") +
         scale_shape_manual(values=c(1,19), name="Body Mass Index") +
         theme_bw() +
         theme(panel.grid=element_blank(), text=element_text(size=8),
               legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
               legend.position="bottom")
autoplot(PCA_scaled_bmi, data=data_scaled_bmi, shape="bmi2") +
         scale_shape_manual(values=c(1,19,15), name="Body Mass Index 2") +
         theme_bw() +
         theme(panel.grid=element_blank(), text=element_text(size=8),
               legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
               legend.position="bottom")
```

### Body Condition 
```{r}
autoplot(PCA_scaled_bc, data=data_scaled_bc, shape="bc") +
         scale_shape_manual(values=c(1,19,15), name="Body Condition") +
         theme_bw() +
         theme(panel.grid=element_blank(), text=element_text(size=8),
               legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
               legend.position="bottom")

PCA_scaled_bc_df <- data.frame(x=PCA_scaled_bc$x[,1],
                               y=PCA_scaled_bc$x[,2],
                               bc=data_scaled_bc$bc)
PCA_scaled_bc_ggplot <- 
  ggplot(PCA_scaled_bc_df, 
         aes(x=x, 
             y=y,
             col=bc)) +
  geom_point() +
  stat_ellipse() +
  labs(x="PC1 (24.89%)",
       y = "PC2 (15.08%)",
       name="Body Condition") +
  scale_color_manual("Body Condition",
                     values=c("#9d9d9d","#008ba2", "#000000")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position.inside=c(0.1, 0.1))

PCA_scaled_bc_ggplot
```

### Body Mass Index
```{r}
PCA_scaled_bmi_df <- data.frame(x=PCA_scaled_bmi$x[,1],
                               y=PCA_scaled_bmi$x[,2],
                               bmi=data_scaled_bmi$bmi)
PCA_scaled_bmi_ggplot <- 
  ggplot(PCA_scaled_bmi_df, 
         aes(x=x, 
             y=y,
             col=bmi)) +
  geom_point() +
  stat_ellipse() +
  labs(x="PC1 (25.11%)",
       y = "PC2 (14.78%)",
       name="Body Mass Index") +
  scale_color_manual("Body Mass Index",
                     values=c("#9d9d9d","#008ba2")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position.inside=c(0.1, 0.1))

PCA_scaled_bmi_ggplot
```

### Body Mass Index 2
```{r}
PCA_scaled_bmi2_df <- data.frame(x=PCA_scaled_bmi$x[,1],
                                 y=PCA_scaled_bmi$x[,2],
                                 bmi2=data_scaled_bmi$bmi2)
PCA_scaled_bmi2_ggplot <- 
  ggplot(PCA_scaled_bmi2_df, 
         aes(x=x, 
             y=y,
             col=bmi2)) +
  geom_point() +
  stat_ellipse() +
  labs(x="PC1 (25.11%)",
       y = "PC2 (14.78%)",
       name="Body Mass Index") +
  scale_color_manual("Body Mass Index",
                     values=c("#9d9d9d","#008ba2", "black")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position.inside=c(0.1, 0.1))

PCA_scaled_bmi2_ggplot
```


## Test for normality
```{r}
# Create dataframe for results
norm_results <- data.frame(matrix(ncol=0, nrow=9)) %>%
  mutate(cyto=colnames(data_scaled)[1:9],
         W=NA, 
         pvalue=NA)

# Truncate data
data_norm <-
  data_scaled[,1:9]

# Test!
for (i in 1:9) {
  test <- shapiro.test(data_norm[,i])
  norm_results[i,2] <- round(test$statistic, digits=3)
  norm_results[i,3] <- test$p.value
}

norm_results # all not normally distributed
```

## Cramer Test 
Multivariate composition of the cytokine pattern between each molt stage pairwise comparison It can handle multiple variables, potentially complex interactions and correlations.  
```{r}
# Separate data by body condition
poor <- 
  data_scaled %>% 
  filter(bc=="Poor") %>% 
  select(1:9) %>% 
  data.matrix()

average <- 
  data_scaled %>% 
  filter(bc=="Average") %>% 
  select(1:9) %>% 
  data.matrix()

robust <- 
  data_scaled %>% 
  filter(bc=="Robust") %>% 
  select(1:9) %>% 
  data.matrix()

# Run Cramer Tests!
cramer.test(poor, average)
cramer.test(average, robust)
cramer.test(poor, robust) 

# Separate data by Body Mass Index
bmi_poor <-
  data_scaled %>% 
  filter(bmi=="Poor") %>% 
  select(1:9) %>% 
  data.matrix()

bmi_robust <- 
  data_scaled %>% 
  filter(bmi=="Robust") %>% 
  select(1:9) %>% 
  data.matrix()

cramer.test(bmi_poor, bmi_robust) 

# Separate data by Body Mass Index2
bmi2_poor <-
  data_scaled %>% 
  filter(bmi2=="Poor") %>% 
  select(1:9) %>% 
  data.matrix()

bmi2_avg <- 
  data_scaled %>% 
  filter(bmi2=="Average") %>% 
  select(1:9) %>% 
  data.matrix()

bmi2_robust <- 
  data_scaled %>% 
  filter(bmi2=="Robust") %>% 
  select(1:9) %>% 
  data.matrix()

# Cramer test
cramer.test(bmi2_poor, bmi2_avg)
cramer.test(bmi2_avg, bmi2_robust)
cramer.test(bmi2_poor, bmi2_robust) 
```

## Cytokine Importance
Not doing because no sig difference among body condition groups. 
```{r, message=FALSE}
# Prepare data for CART
data_cart <- 
  lapply(data_scaled, c) %>% 
  data.frame() %>% 
  select(!sample) %>% 
  mutate(bc=factor(bc, levels=c("Poor", "Average", "Robust"))) 

data_cart_bmi <-
  lapply(data_scaled, c) %>% 
  data.frame() %>% 
  select(!sample) %>% 
  mutate(bmi=factor(bmi, levels=c("Poor", "Robust")))

data_cart_bmi2 <-
  lapply(data_scaled, c) %>% 
  data.frame() %>% 
  select(!sample) %>% 
  mutate(bmi2=factor(bmi2, levels=c("Poor", "Average", "Robust")))
```

### Cross Validation
Not doing because no sig difference among body condition groups. 

## How different groups are

### PLSDA - Body Condition
```{r, warning=FALSE}
pls_x <-
  data_cart %>% 
  filter(!bc=="NA") %>% 
  select(1:9)

pls_y <- 
  data_cart %>% 
  select(bc) %>% 
  filter(!bc=="NA")

# PLSDA Model
plsda_model <- mixOmics::plsda(pls_x,
                               pls_y$bc,
                               scale=FALSE,
                               ncomp=9)

plotIndiv(plsda_model,
          ellipse=TRUE, 
          ind.names=FALSE,
          col.per.group=c("#9d9d9d","#008ba2", "black"),
          pch=16,
          legend=TRUE, 
          legend.title="Body Condition",
          style="lattice",
          title="PLS-DA Comp 1 - 2")

biplot(plsda_model, ind.names=FALSE, legend.title="Body Condition")

# ggplot2
plsda_ggplot <- 
  data.frame(x = plsda_model$variates$X[,1], 
             y = plsda_model$variates$X[,2],
             bc = plsda_model$Y)

bc_plsda_ggplot <- 
  ggplot(plsda_ggplot, 
         aes(x=x, 
             y=y,
             col=bc)) +
  geom_point() +
  stat_ellipse() +
  labs(x="x-variate 1 (21%)",
       y = "x-variate 2 (9%)",
       name="Body Condition") +
  scale_color_manual("Body Condition",
                     values=c("#9d9d9d","#008ba2", "black")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position=c(0.1, 0.1))

bc_plsda_ggplot
```

### PLSDA - Body Mass Index
```{r, warning=FALSE}
pls_x_bmi <-
  data_cart_bmi %>% 
  filter(!bmi=="NA") %>% 
  select(1:9)

pls_y_bmi <- 
  data_cart_bmi %>% 
  filter(!bmi=="NA") %>% 
  select(bmi)

# PLSDA Model
plsda_model_bmi <- mixOmics::plsda(pls_x_bmi,
                               pls_y_bmi$bmi,
                               scale=FALSE,
                               ncomp=9)

plotIndiv(plsda_model_bmi,
          ellipse=TRUE, 
          ind.names=FALSE,
          col.per.group=c("#9d9d9d","#008ba2"),
          pch=16,
          legend=TRUE, 
          legend.title="Body Mass Index",
          style="lattice",
          title="PLS-DA Comp 1 - 2")

biplot(plsda_model_bmi, ind.names=FALSE, legend.title="Body Mass Index")

# ggplot2
plsda_ggplot_bmi <- 
  data.frame(x = plsda_model_bmi$variates$X[,1], 
             y = plsda_model_bmi$variates$X[,2],
             bmi = plsda_model_bmi$Y)

bmi_plsda_ggplot <- 
  ggplot(plsda_ggplot_bmi, 
         aes(x=x, 
             y=y,
             col=bmi)) +
  geom_point() +
  stat_ellipse() +
  labs(x="x-variate 1 (22%)",
       y = "x-variate 2 (15%)",
       name="Body Mass Index") +
  scale_color_manual("Body Mass Index",
                     values=c("#9d9d9d","#008ba2")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position=c(0.1, 0.1))

bmi_plsda_ggplot
```

### PLSDA - Body Mass Index 2
```{r, warning=FALSE}
pls_x_bmi2 <-
  data_cart_bmi %>% 
  filter(!bmi2=="NA") %>% 
  select(1:9)

pls_y_bmi2 <- 
  data_cart_bmi %>% 
  filter(!bmi2=="NA") %>% 
  select(bmi2)

# PLSDA Model
plsda_model_bmi2 <- mixOmics::plsda(pls_x_bmi2,
                                    pls_y_bmi2$bmi2,
                                    scale=FALSE,
                                    ncomp=9)

plotIndiv(plsda_model_bmi2,
          ellipse=TRUE, 
          ind.names=FALSE,
          col.per.group=c("#9d9d9d","#008ba2", "black"),
          pch=16,
          legend=TRUE, 
          legend.title="Body Mass Index 2",
          style="lattice",
          title="PLS-DA Comp 1 - 2")

biplot(plsda_model_bmi2, ind.names=FALSE, legend.title="Body Mass Index 2")

# ggplot2
plsda_ggplot_bmi2 <- 
  data.frame(x = plsda_model_bmi2$variates$X[,1], 
             y = plsda_model_bmi2$variates$X[,2],
             bmi2 = plsda_model_bmi2$Y)

bmi2_plsda_ggplot <- 
  ggplot(plsda_ggplot_bmi2, 
         aes(x=x, 
             y=y,
             col=bmi2)) +
  geom_point() +
  stat_ellipse() +
  labs(x="x-variate 1 (20%)",
       y = "x-variate 2 (9%)",
       name="Body Mass Index 2") +
  scale_color_manual("Body Mass Index 2",
                     values=c("#9d9d9d","#008ba2", "black")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position=c(0.1, 0.1))

bmi2_plsda_ggplot
```
