---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Body condition ~ cytokines

## Libraries
```{r, warning=FALSE, message=FALSE}
library(vegan)
library(mixOmics)
library(tidyverse)
library(ggfortify)
library(cramer)
library(mlr3)
library(paradox)
library(mlr3tuning)
library(rpart)
library(rpart.plot)
library(pairwiseAdonis)
library(ggmosaic)
```

## Data
```{r, message=FALSE, warning=FALSE}
cyto <- read.csv("Output Files/cleaned_cytokine.csv")
cytobin <- read.csv("Output Files/cleaned_cytokine_bin.csv")
cytoscale <- read.csv("Output Files/cleaned_cytokine_scaled.csv")
meta <- read.csv("Output Files/metadata_cyto.csv")

# Merge cytokine & metadata 
data <- merge(cyto, meta, by="Sample") %>% 
  filter(!is.na(bc_bin))

databin <- 
  merge(cytobin, meta, by="Sample") %>% 
  filter(!is.na(bc_bin))

datascale <- 
  merge(cytoscale, meta, by="Sample") %>% 
  filter(!is.na(bc_bin))
```


## Body Condition Range & Bins
```{r}
quant <- quantile(meta$bodcond, na.rm=TRUE)

ggplot(meta, aes(x=bodcond)) +
  geom_histogram(bins=10, color="black", fill="gray80") +
  labs(x="Body Condition Index", y="Frequency") +
  geom_vline(xintercept=c(quant[2], quant[3], quant[4]), colour="red", size=1) +
  theme_classic()
```

## PCA
```{r, warning=FALSE}
PCA_scaled_bc <- 
  prcomp(~IFNg+IL.2+IL.6+IL.7+IL.8+IL.10+IL.15+IL.18+KC.like, 
         data=datascale,
         center=FALSE, 
         scale=FALSE)
```

### Visualize
```{r}
PCA_scaled_bc_df <- data.frame(x=PCA_scaled_bc$x[,1],
                               y=PCA_scaled_bc$x[,2],
                               bc=datascale$bc)
pca_labelx <-
  paste("PC1 (", 
                round(abs(summary(PCA_scaled_bc)$importance[2,1]*100), digits=2),
                "%)",
                sep="")
pca_labely <-
  paste("PC2 (", 
                round(abs(summary(PCA_scaled_bc)$importance[2,2]*100), digits=2),
                "%)",
                sep="")

PCA_scaled_bc_ggplot <- 
  ggplot(PCA_scaled_bc_df, 
         aes(x=x, 
             y=y,
             col=bc)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely,
       name="Body Condition") +
  scale_color_manual("Body Condition",
                     values=c("#9d9d9d","#008ba2", "#000000")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position.inside=c(0.1, 0.1))

PCA_scaled_bc_ggplot
```


## PERMANOVA

### Cytokine Presence/Absence

#### Create similarity matrix (Sorensen)
```{r}
databin_perm <- 
  databin %>% 
  select(2:10)

databin_dist <- 
  betadiver(databin_perm, method=11)

plot(databin_dist)

# Dissimilarity summary stats
mean(1-databin_dist); min(1-databin_dist); max(1-databin_dist); median(1-databin_dist)
```

#### PERMANOVA with dissimilarity (1-Sorensen)
R2 = partial R2 = factor (molt stage) is explaining at least R2 * 100% of variation in variables (cytokines)
```{r}
databin_adonis <- 
  adonis2((1-databin_dist) ~ bc_bin, data=databin, permutations=4999)
databin_adonis

# R2
databin_adonis$R2[1]*100
```

#### Visualizing multivariate groups

##### Principal coordinates analysis 
```{r, warning=FALSE}
# Use 1 - Sorensen's for dissimilarity 
databin_pco <- cmdscale(1-databin_dist, eig=TRUE)
```

##### Plot
```{r}
plot(databin_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)

ordispider(jitter(databin_pco$points, factor=1), factor(databin$bc_bin), show.groups="Poor", 
           col="deepskyblue", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(databin_pco$points, factor=1), factor(databin$bc_bin), show.groups="Average", 
           col="firebrick", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(databin_pco$points, factor=1), factor(databin$bc_bin), show.groups="Robust", 
           col="darkmagenta", label = TRUE, display="sites", lwd=1.5, cex=1.2)
```

#### Homogeneity of group dispersions
```{r}
disper <- 
  betadisper(databin_dist, group=databin$bc_bin, type="centroid")
anova(disper)
TukeyHSD(disper)
disper
boxplot(disper)
```

#### Posthoc pairwise tests
None - no sig difference in multivariate distributions


### Cytokine Concentration
```{r}
data_perm <-
  data %>% 
  select(2:10)

data_dist <- 
  vegdist(data_perm, method="bray")

data_adonis <- adonis2(data_dist ~ bc_bin, data=data, 
                       permuations=4999)
data_adonis

# R2
data_adonis$R2[1]*100
```

#### Visualizing multivariate groups

##### Principal coordinates analysis 
```{r, warning=FALSE}
# Use Bray-Curtis dissimilarity matrix
data_pco <- cmdscale(data_dist, eig=TRUE)
```

##### Plot
```{r}
plot(data_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)

ordispider(jitter(data_pco$points, factor=1), factor(data$bc_bin), show.groups="Poor", 
           col="deepskyblue", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(data_pco$points, factor=1), factor(data$bc_bin), show.groups="Average", 
           col="firebrick", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(data_pco$points, factor=1), factor(data$bc_bin), show.groups="Robust", 
           col="darkmagenta", label = TRUE, display="sites", lwd=1.5, cex=1.2)
```

#### Homogeneity of group dispersions
```{r}
betadisper(data_dist, group=data$bc, type="centroid")

disper <- 
  betadisper(data_dist, group=data$bc_bin, type="centroid")
anova(disper)
disper
boxplot(disper)
```

#### Posthoc pairwise tests
None - no sig difference in multivariate distributions


## Cramer Test 
Multivariate composition of the cytokine pattern between each molt stage pairwise comparison It can handle multiple variables, potentially complex interactions and correlations.  
```{r}
# Separate data by body condition
poor <- 
  datascale %>% 
  filter(bc_bin=="Poor") %>% 
  select(2:10) %>% 
  data.matrix()

average <- 
  datascale %>% 
  filter(bc_bin=="Average") %>% 
  select(2:10) %>% 
  data.matrix()

robust <- 
  datascale %>% 
  filter(bc_bin=="Robust") %>% 
  select(2:10) %>% 
  data.matrix()

# Run Cramer Tests!
cramer.test(poor, average)
cramer.test(average, robust)
cramer.test(poor, robust) 
```

## Cytokine Importance

### Prepare data for CART
```{r, message=FALSE}
data_cart <- 
  datascale %>% 
  select(2:10, bc_bin) %>% 
  mutate(bc_bin=factor(bc_bin, levels=c("Poor", "Average", "Robust")))
```

### Cross Validation
See reference: https://mlr3book.mlr-org.com/
```{r}
# Creating task and learner
task <- as_task_classif(bc_bin ~ ., 
                        data = data_cart)

task <- task$set_col_roles(cols="bc_bin",
                           add_to="stratum")

learner <- lrn("classif.rpart",
               predict_type = "prob",
               maxdepth = to_tune(2, 5),
               minbucket = to_tune(1, 25),
               minsplit = to_tune(1, 30)
               )

# Define tuning instance - info ~ tuning process
instance <- ti(task = task,
               learner = learner,
               resampling = rsmp("cv", folds = 10),
               measures = msr("classif.ce"),
               terminator = trm("none")
               )

# Define how to tune the model
tuner <- tnr("grid_search", 
             batch_size = 10
             )

# Trigger the tuning process
#tuner$optimize(instance)
 
# optimal values: 
  # maxdepth = 5
  # minbucket = 6
  # minsplit = 10
  # classif.ce = 0.5360839
```

### Run model with optimized parameters
```{r}
cart_model <- rpart(formula=bc_bin ~ .,
                    data=data_cart,
                    method="class",
                    maxdepth=5,
                    minbucket=6,
                    minsplit=10)

# plot optimized tree
rpart.plot(cart_model)
cart_model$variable.importance 
  # IFNg, IL-18 are two most important cytokines
```

### Plot important cytokines
#### Variable Importance
```{r}
varimp <- 
  data.frame(imp=cart_model$variable.importance) %>% 
  rownames_to_column() %>% 
  rename("variable" = rowname) %>% 
  arrange(imp) %>%
  mutate(variable=gsub("\\.","-",variable),
         variable = forcats::fct_inorder(variable))

varimp_plot <- 
  ggplot(varimp) + 
  geom_segment(aes(x = variable, 
                   y = 0, 
                   xend = variable, 
                   yend = imp), 
               linewidth = 0.5, 
               alpha = 0.7) +
  geom_point(aes(x = variable, 
                 y = imp), 
             size = 1, 
             show.legend = F,
             col="black") +
  labs(x="Cytokine", 
       y="Variable Importance") +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=10, color="black"))
varimp_plot
```

#### Proportions
```{r}
prop.table(table(databin$bc_bin, databin$IFNg), margin=1)*100

ifng_prop <- 
  databin %>% 
  mutate(IFNg=gsub("1", "P", IFNg),
         IFNg=gsub("0", "A", IFNg),
         IFNg=factor(IFNg, levels=c("P", "A"))) %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IFNg, bc_bin), fill=IFNg)) +
  scale_fill_manual(values=c("P"="gray60", "A"="lightgray")) +
  labs(y="Presence/Absence") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none",
        text = element_text(size=10),
        plot.title = element_text(hjust = 0.5))
ifng_prop

il18_prop <-
  databin %>% 
  mutate(IL.18=gsub("1", "P", IL.18),
         IL.18=gsub("0", "A", IL.18),
         IL.18=factor(IL.18, levels=c("P", "A"))) %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.18, bc_bin), fill=IL.18)) +
  scale_fill_manual(values=c("P"="gray60", "A"="lightgray")) +
  labs(y="Presence/Absence") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none",
        text = element_text(size=10),
        plot.title = element_text(hjust = 0.5))
il18_prop
```

#### Concentration
```{r}
ifng <-
  data %>% 
  select(bc_bin, IFNg) %>% 
  filter(!IFNg==0) %>% 
  ggplot(., aes(x=bc_bin, y=log(IFNg))) +
  geom_boxplot(color="black", fill="gray60", lwd=0.25,
               outliers=FALSE) +
  geom_point(size=0.75, position=position_jitter(width=0.2)) +
  labs(x=NULL, y="log(Concentration pg/mL)") +
  stat_n_text(size=3) +
  theme_classic() +
  theme(text = element_text(size=10),
         panel.border=element_rect(color="black", fill=NA, linewidth=0.5))
ifng  

il18 <-
  data %>% 
  select(bc_bin, IL.18) %>% 
  filter(!IL.18==0) %>% 
  ggplot(., aes(x=bc_bin, y=log(IL.18))) +
  geom_boxplot(color="black", fill="gray60", lwd=0.25,
               outliers=FALSE) +
  geom_point(size=0.75, position=position_jitter(width=0.2)) +
  labs(x=NULL, y="log(Concentration pg/mL)") +
  stat_n_text(size=3) +
  theme_classic() +
  theme(text = element_text(size=10),
         panel.border=element_rect(color="black", fill=NA, linewidth=0.5))
il18
```

#### Combine
```{r}
left <- grid.arrange(ifng_prop, ifng, ncol=1,
                     top = textGrob("IFNg", gp=gpar(fontsize=10)))
right <- grid.arrange(il18_prop, il18, ncol=1,
                      top = textGrob("IL-18", gp=gpar(fontsize=10)))

jpeg("Figures/cytoimportance_bodycondition.jpeg", width=7, height=8, units="in", res=300)
all <- grid.arrange(varimp_plot, left, right, 
                    layout_matrix=rbind(c(1, 1),
                                        c(2, 3)),
                    heights=c(1,2),
                    bottom=textGrob("Body Condition", gp=gpar(fontsize=10)))
grid.polygon(x=c(0, 0, 1, 1),
             y=c(0, 0.68, 0.68, 0),
             gp=gpar(fill=NA))
grid.polygon(x=c(0.02, 0.02, 0.97, 0.97),
             y=c(0.93, 0.99, 0.99, 0.93),
             gp=gpar(fill=NA))
dev.off()
```


## How different groups are

### PLSDA - Body Condition
#### Prepare data
```{r, warning=FALSE}
pls_x <-
  data_cart[,1:9]

pls_y <- 
  data_cart %>% 
  select(bc_bin) 
```

#### PLSDA Model
```{r}
plsda_model <- mixOmics::plsda(pls_x,
                               pls_y$bc_bin,
                               scale=TRUE,
                               ncomp=9)
```

#### Visualizations
```{r}
# Set x and y axis labels based on model loadings
labelx <- paste("X-variate 1: ", 
                round(abs(plsda_model$prop_expl_var$X[1]*100), digits=0),
                "% expl. var",
                sep="")

labely <- paste("X-variate 2: ", 
                round(abs(plsda_model$prop_expl_var$X[2]*100), digits=0),
                "% expl. var",
                sep="")

# ggplot2
plsda_ggplot <- 
  data.frame(x = plsda_model$variates$X[,1], 
             y = plsda_model$variates$X[,2],
             bc_bin = plsda_model$Y)

bc_plsda_ggplot <- 
  ggplot(plsda_ggplot, 
         aes(x=x, 
             y=y,
             col=bc_bin)) +
  geom_point() +
  stat_ellipse() +
  labs(x=labelx,
       y = labely,
       name="Body Condition") +
  scale_color_manual("Body Condition",
                     values=c("#9d9d9d","#008ba2", "black")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position=c(0.1, 0.1))

bc_plsda_ggplot
```

#### Cytokine Contributions
```{r}
plsda_loadings <- 
  plsda_model$loadings$X %>% 
  data.frame() %>% 
  select(1:2) %>% 
  arrange(comp1)
plsda_loadings

biplot(plsda_model, ind.names=FALSE, legend.title="Body Condition")

plotVar(plsda_model)
```
