---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# IAV ~ cytokines

## Libraries
```{r, warning=FALSE, message=FALSE}
library(ggfortify)
library(ggpubr)
library(gridExtra)
library(ggmosaic)
library(cramer)
library(mlr3)
library(paradox)
library(mlr3tuning)
library(rpart)
library(rpart.plot)
library(mixOmics)
library(tidyverse)
library(fasano.franceschini.test)
library(EnvStats)
library(grid)
```

## Data
```{r, message=FALSE, warning=FALSE}
cyto <- read.csv("Output Files/cleaned_cytokine.csv")
meta <- read.csv("Output Files/metadata_cyto.csv")
mindc <- read.csv("Input Files/mindc.csv") %>% 
  mutate(half=value/2)
cytobin <- read.csv("Output Files/cleaned_cytokine_bin.csv")

# Merge cytokine & metadata 
data <- merge(cyto, meta, by="Sample")
databin <- merge(cytobin, meta, by="Sample")

# Get rid of variables that are not of interest & cytokines with < 10% (11) detections
data_tidy <-
  data %>%
  select(2:14) %>% 
  select_if(function(x) sum(x>0) > 11) %>% 
  mutate(iav=data$iav, 
         molt.stage=data$molt.stage,
         bc_bin=data$bc_bin,
         location=data$location,
         stage=factor(data$stage, levels=c("control", "acute", "peak", "late")),
         sample=data$Sample)
```

## PCA on raw cytokine data
### Infection
```{r, warning=FALSE}
PCA_iav <- prcomp(data_tidy[,1:9], center=FALSE, scale=FALSE)

raw_iav <- autoplot(PCA_iav, data=data_tidy, shape="iav") +
           scale_shape_manual(values=c(1,19), name="Influenza A Virus", 
                              labels=c("Negative", "Positive")) +
           theme_bw() +
           theme(panel.grid=element_blank(), text=element_text(size=8),
                legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
                 legend.position="bottom")

iav_legend<-get_legend(raw_iav)
legend<-as_ggplot(iav_legend)

raw_iav <- 
  raw_iav + 
  theme(legend.position="none")
```

### Molt Stage
```{r}
data_raw_molt <- 
  data_tidy %>% 
  filter(!is.na(molt.stage))

PCA_molt <- prcomp(data_raw_molt[,1:9], center=FALSE, scale=FALSE)

raw_molt <- autoplot(PCA_molt, data=data_raw_molt, shape="molt.stage") +
          scale_shape_manual(values=c(0,16,8,2), name="Molt Stage") +
          theme_bw() +
          theme(panel.grid=element_blank(), text=element_text(size=8),
                legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
                legend.position="bottom")

molt_legend<-get_legend(raw_molt)
legend<-as_ggplot(molt_legend)

raw_molt <- 
  raw_molt + 
  theme(legend.position="none")
```

### Body Condition
```{r}
data_raw_bc <- 
  data_tidy %>% 
  filter(!is.na(bc_bin))

PCA_bc <- prcomp(data_raw_bc[,1:9], center=FALSE, scale=FALSE)

raw_bc <- autoplot(PCA_bc, data=data_raw_bc, shape="bc_bin") +
          scale_shape_manual(values=c(0,16,8), name="Body Condition",
                             labels=c("Low", "Mid", "High")) +
          theme_bw() +
          theme(panel.grid=element_blank(), text=element_text(size=8),
                legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
                legend.position="bottom")

bc_legend<-get_legend(raw_bc)
legend<-as_ggplot(bc_legend)

raw_bc <- 
  raw_bc + 
  theme(legend.position="none")
```

### Location
```{r}
PCA_loc <- prcomp(data_tidy[,1:9], center=FALSE, scale=FALSE)

raw_loc <- autoplot(PCA_loc, data=data_tidy, shape="location") + 
          scale_shape_manual(values=c(0,16,8), name="Location") +
          theme_bw() +
          theme(panel.grid=element_blank(), text=element_text(size=8),
                legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
                legend.position="bottom")

loc_legend<-get_legend(raw_loc)
legend<-as_ggplot(loc_legend)

raw_loc <- 
  raw_loc + 
  theme(legend.position="none")
```

## PCA on log-transformed, scaled, centered cytokine data
### Prepare data
```{r}
# Make all non-detects half the limit of detection
data_half <- data_tidy
for (i in 1:nrow(data_half)) {
  for (j in 1:9) {
     {if (data_half[i,j]==0) {
       k <- which(mindc==colnames(data_half)[j])
       data_half[i,j] <- mindc[k,3]
     }
       else{next}
       }
  }
}

# Log transform, scale, and center data
data_log <- 
  data_half %>% 
  mutate_at(1:9, log)

data_scaled <- 
  data_half %>% 
  mutate_at(1:9, log) %>% 
  mutate_at(1:9, function(x) scale(x, center=TRUE, scale=TRUE))
```

### Infection Stage
```{r}
PCA_scaled_iav <- prcomp(data_scaled[,1:9], center=FALSE, scale=FALSE)

scaled_iav <- autoplot(PCA_scaled_iav, data=data_scaled, shape="iav") +
              scale_shape_manual(values=c(1, 19), name="Influenza A Virus", 
                                 labels=c("Negative", "Positive")) +
              theme_bw() +
              theme(panel.grid=element_blank(), text=element_text(size=8),
                    legend.background=element_rect(size=0.5, linetype="solid", 
                                                   colour ="black"),
                    legend.position="none")
```

### Molt Stage
```{r}
data_scaled_molt <- 
  data_scaled %>% 
  filter(!is.na(molt.stage))

PCA_scaled_molt <- prcomp(data_scaled_molt[,1:9], center=FALSE, scale=FALSE)

scaled_molt <- autoplot(PCA_scaled_molt, data=data_scaled_molt, shape="molt.stage") +
              scale_shape_manual(values=c(0,16,8,2), name="Molt Stage") +
              theme_bw() +
              theme(panel.grid=element_blank(), text=element_text(size=8),
                    legend.background=element_rect(size=0.5, linetype="solid", 
                                                   colour ="black"),
                    legend.position="none")
```

### Body Condition
```{r}
data_scaled_bc <- 
  data_scaled %>% 
  filter(!is.na(bc_bin))

PCA_scaled_bc <- prcomp(data_scaled_bc[,1:9], center=FALSE, scale=FALSE)

scaled_bc <- autoplot(PCA_scaled_bc, data=data_scaled_bc, shape="bc_bin") +
             scale_shape_manual(values=c(0,16,8), name="Body Condition") +
             theme_bw() +
             theme(panel.grid=element_blank(), text=element_text(size=8),
                   legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
                   legend.position="none")
```

#### Location
```{r}
PCA_scaled_loc <- prcomp(data_scaled[,1:9], center=FALSE, scale=FALSE)

scaled_loc <- autoplot(PCA_scaled_loc, data=data_scaled, shape="location") +
          scale_shape_manual(values=c(0,16,8), name="Location") +
          theme_bw() +
          theme(panel.grid=element_blank(), text=element_text(size=8),
                legend.background=element_rect(size=0.5, linetype="solid", 
                                               colour ="black"),
                legend.position="none")
```

### Combine
```{r}
cyto_pca_iav <- grid.arrange(arrangeGrob(raw_iav, scaled_iav, ncol=2),
                             arrangeGrob(iav_legend, ncol=1),
                             heights=c(5,1))

cyto_pcas <- grid.arrange(arrangeGrob(scaled_iav, scaled_molt, ncol=2), 
                          arrangeGrob(iav_legend, molt_legend, ncol=2), 
                          arrangeGrob(scaled_bc, scaled_loc, ncol=2),
                          arrangeGrob(bc_legend, loc_legend, ncol=2),
                          heights=c(5,1,5,1))
```

## Test for normality
```{r}
# Create dataframe for results
norm_results <- data.frame(matrix(ncol=0, nrow=9)) %>%
  mutate(cyto=colnames(data_scaled)[1:9],
         W=NA, 
         pvalue=NA)

# Truncate data
data_norm <-
  data_scaled[,1:9]

# Test!
for (i in 1:9) {
  test <- shapiro.test(data_norm[,i])
  norm_results[i,2] <- round(test$statistic, digits=3)
  norm_results[i,3] <- test$p.value
}

norm_results # all not normally distributed
```

## Cramer Test 
Multivariate composition of the cytokine pattern between each infection stage group with its control. It can handle multiple variables, potentially complex interactions and correlations.  
```{r}
# Separate data by infection stage
infected <- 
  data_scaled %>% 
  filter(iav=="pos") %>% 
  select(1:9) %>% 
  data.matrix()

uninfected <- 
  data_scaled %>% 
  filter(iav=="neg") %>% 
  select(1:9) %>% 
  data.matrix()

control <- 
  data_scaled %>% 
  filter(stage=="control") %>% 
  select(1:9) %>% 
  mutate_at(1:9, as.numeric) %>% 
  data.matrix()

acute <- 
  data_scaled %>% 
  filter(stage=="acute") %>% 
  select(1:9) %>% 
  mutate_at(1:9, as.numeric) %>% 
  data.matrix()

peak <-
  data_scaled %>% 
  filter(stage=="peak") %>% 
  select(1:9) %>% 
  mutate_at(1:9, as.numeric) %>% 
  data.matrix()

late <- 
  data_scaled %>% 
  filter(stage=="late") %>% 
  select(1:9) %>% 
  mutate_at(1:9, as.numeric) %>% 
  data.matrix()

# Run Cramer Tests!
cramer.test(infected, uninfected)

cramer.test(control, acute)
cramer.test(control, peak)
cramer.test(control, late)
```

## Cytokine Importance
```{r, message=FALSE}
# Prepare data for CART
data_cart <- 
  lapply(data_scaled, c) %>% 
  data.frame() %>% 
  select(!location & !molt.stage & !bc_bin & !stage & !sample) %>% 
  mutate(iav=factor(iav))
```

### Cross Validation
```{r}
# Creating task and learner
task <- as_task_classif(iav ~ ., 
                        data = data_cart)
        
task <- task$set_col_roles(cols="iav",
                           add_to="stratum")

learner <- lrn("classif.rpart",
               predict_type = "prob",
               maxdepth = to_tune(2, 5),
               minbucket = to_tune(1, 30),
               minsplit = to_tune(1, 30)
               )

# Define tuning instance - info ~ tuning process
instance <- ti(task = task,
               learner = learner,
               resampling = rsmp("cv", folds = 10),
               measures = msr("classif.ce"),
               terminator = trm("none")
               )

# Define how to tune the model
tuner <- tnr("grid_search", 
             batch_size = 10
             )

# Trigger the tuning process
set.seed(7)
# tuner$optimize(instance)

# optimal values: 
  # maxdepth = 5
  # minbucket = 20
  # minsplit = 20
  # classif.ce = 0.2757576

# Model with optimized parameters
cart_model <- rpart(formula=iav ~ .,
                    data=data_cart,
                    method="class",
                    maxdepth=5,
                    minbucket=20,
                    minsplit=20)

# plot optimized tree
rpart.plot(cart_model)
printcp(cart_model)
cart_model$variable.importance 
  # IL-6 and IL-10 are top two cytokines
```

### Manuscript Plot
Variable importance: the sum of the goodness of split measures for each split for which it was the primary variable, plus goodness * (adjusted agreement) for all splits in which it was a surrogate. In the printout these are scaled to sum to 100 and the rounded values are shown, omitting any variable whose proportion is less than 1%. (https://cran.r-project.org/web/packages/rpart/vignettes/longintro.pdf)

#### Variable Importance
```{r}
varimp <- 
  data.frame(imp=cart_model$variable.importance) %>% 
  rownames_to_column() %>% 
  rename("variable" = rowname) %>% 
  arrange(imp) %>%
  mutate(variable=gsub("\\.","-",variable),
         variable = forcats::fct_inorder(variable))

varimp_plot <- 
  ggplot(varimp) + 
  geom_segment(aes(x = variable, 
                   y = 0, 
                   xend = variable, 
                   yend = imp), 
               linewidth = 0.5, 
               alpha = 0.7) +
  geom_point(aes(x = variable, 
                 y = imp), 
             size = 1, 
             show.legend = F,
             col="black") +
  labs(x="Cytokine", 
       y="Variable Importance") +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=10, color="black"))
```

#### Data Prep
```{r}
databin <-
  databin %>% 
  mutate(iav=gsub("neg", "Neg", iav),
         iav=gsub("pos", "Pos", iav))

data <-
  data %>% 
  mutate(iav=gsub("neg", "Neg", iav),
         iav=gsub("pos", "Pos", iav))
```

#### Proportions
```{r}
il6_prop <- 
  databin %>%
  mutate(IL.6=gsub("1", "P", IL.6),
         IL.6=gsub("0", "A", IL.6),
         IL.6=factor(IL.6, levels=c("P", "A"))) %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.6,iav), fill=IL.6)) +
  scale_fill_manual(values=c("P"="gray60", "A"="lightgray")) +
  labs(y="IL-6") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(), 
        axis.text.x= element_blank(),
        legend.position="none",
        text = element_text(size = 10))

il10_prop <- 
  databin %>%
  mutate(IL.10=gsub("1", "P", IL.10),
         IL.10=gsub("0", "A", IL.10),
         IL.10=factor(IL.10, levels=c("P", "A"))) %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.10,iav), fill=IL.10)) +
  scale_fill_manual(values=c("P"="gray60", "A"="lightgray")) +
  labs(y="IL-10") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position="none",
        text = element_text(size = 10))

kc_prop <- 
  databin %>%
  mutate(KC.like=gsub("1", "P", KC.like),
         KC.like=gsub("0", "A", KC.like),
         KC.like=factor(KC.like, levels=c("P", "A"))) %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(KC.like,iav), fill=KC.like)) +
  scale_fill_manual(values=c("P"="gray60", "A"="lightgray")) +
  labs(y="KC-like") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none",
        text = element_text(size = 10))
```


#### Concentrations - Raw
```{r}
il6 <- 
  data %>% 
  select(iav, IL.6) %>% 
  filter(!IL.6 == 0) %>% 
  ggplot(., aes(x=iav, y=log(IL.6))) +
  geom_boxplot(color="gray60", fill="gray60", lwd=0.25) +
  geom_point(size=0.5) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=1.5) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        text = element_text(size = 10))
dat <- ggplot_build(il6)$data[[1]]
il6 <- 
  il6 +
  geom_segment(data=dat, aes(x=xmin, xend=xmax, 
                             y=middle, yend=middle), colour="lightgray")

il6

il10 <- 
  data %>% 
  select(iav, IL.10) %>% 
  filter(!IL.10 == 0) %>% 
  ggplot(., aes(x=iav, y=log(IL.10))) +
  geom_boxplot(color="gray60", fill="gray60", lwd=0.25) +
  geom_point(size=0.5) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=1.5) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        text = element_text(size = 10))
dat <- ggplot_build(il10)$data[[1]]
il10 <-
  il10 +
  geom_segment(data=dat, aes(x=xmin, xend=xmax, 
                             y=middle, yend=middle), colour="lightgray")

kclike <- 
  data %>% 
  select(iav, KC.like) %>% 
  filter(!KC.like==0) %>% 
  ggplot(., aes(x=iav, y=KC.like)) +
  geom_boxplot(color="gray60", fill="gray60", lwd=0.25) +
  geom_point(size=0.5) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=1.5) +
  theme_classic() +
  theme(text = element_text(size = 10))
dat <- ggplot_build(kclike)$data[[1]]
kclike <- 
  kclike +
  geom_segment(data=dat, aes(x=xmin, xend=xmax, 
                             y=middle, yend=middle), colour="lightgray")
```

#### Combine together - Raw
```{r}
prop_grid <- grid.arrange(il6_prop, il10_prop, kc_prop, ncol=1)
conc_grid <- grid.arrange(il6, il10, kclike, ncol=1, 
                          left=textGrob("log(Concentration (pg/mL))", 
                                        rot=90, gp=gpar(fontsize=10)))

cyto_grid <- grid.arrange(prop_grid, conc_grid,
                           ncol=2, 
                           bottom=textGrob("IAV Status", gp=gpar(fontsize=10)))

final_grid <- grid.arrange(varimp_plot, cyto_grid, 
                           ncol=2, widths=c(5.5, 4.5))

ggsave(filename="Figures/cytoimportance.jpg", final_grid, height=5, width=8, units="in")
```


#### Concentrations - Scaled
```{r}
il6_scaled <- 
  data_cart %>% 
  select(iav, IL.6) %>% 
  ggplot(., aes(x=iav, y=IL.6)) +
  geom_boxplot(color="gray60", fill="gray60", lwd=0.25) +
  geom_point(size=0.5) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=1.5) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        text = element_text(size = 10))
dat <- ggplot_build(il6_scaled)$data[[1]]
il6_scaled <- 
  il6_scaled +
  geom_segment(data=dat, aes(x=xmin, xend=xmax, 
                             y=middle, yend=middle), colour="lightgray")

il6_scaled

il10_scaled <- 
  data_cart %>% 
  select(iav, IL.10) %>% 
  ggplot(., aes(x=iav, y=IL.10)) +
  geom_boxplot(color="gray60", fill="gray60", lwd=0.25) +
  geom_point(size=0.5) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=1.5) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        text = element_text(size = 10))
dat <- ggplot_build(il10_scaled)$data[[1]]
il10_scaled <-
  il10_scaled +
  geom_segment(data=dat, aes(x=xmin, xend=xmax, 
                             y=middle, yend=middle), colour="lightgray")
il10_scaled

kclike_scaled <- 
  data_cart %>% 
  select(iav, KC.like) %>% 
  ggplot(., aes(x=iav, y=KC.like)) +
  geom_boxplot(color="gray60", fill="gray60", lwd=0.25) +
  geom_point(size=0.5) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=1.5) +
  theme_classic() +
  theme(text = element_text(size = 10))
dat <- ggplot_build(kclike_scaled)$data[[1]]
kclike_scaled <- 
  kclike_scaled +
  geom_segment(data=dat, aes(x=xmin, xend=xmax, 
                             y=middle, yend=middle), colour="lightgray")
kclike_scaled
```

#### Combine together - Scaled
```{r}
prop_grid <- grid.arrange(il6_prop, il10_prop, kc_prop, ncol=1)
conc_grid <- grid.arrange(il6_scaled, il10_scaled, kclike_scaled, ncol=1, 
                          left=textGrob("Concentration", 
                                        rot=90, gp=gpar(fontsize=10)))

cyto_grid <- grid.arrange(prop_grid, conc_grid,
                           ncol=2, 
                           bottom=textGrob("IAV Status", gp=gpar(fontsize=10)))

final_grid <- grid.arrange(varimp_plot, cyto_grid, 
                           ncol=2, widths=c(5.5, 4.5))

ggsave(filename="Figures/cytoimportance_scaled.jpg", final_grid, height=5, width=8, units="in")
```


## How different IAV+ and IAV- groups are

### PLSDA 

#### Prepare data
```{r}
pls_x <-
 data_cart[,1:9] 

pls_y <- 
  data_cart %>% 
  mutate(iav=gsub("neg", "IAV-", iav),
         iav=gsub("pos", "IAV+", iav),
         iav=as.factor(iav)) %>% 
  select(iav)
```

#### PLSDA Model
```{r}
plsda_model <- mixOmics::plsda(pls_x,
                               pls_y$iav,
                               scale=FALSE,
                               ncomp=9)
```

#### Visualizations
```{r}
plotIndiv(plsda_model,
          ellipse=TRUE, 
          ind.names=FALSE,
          col.per.group=c("#9d9d9d","#008ba2"),
          pch=16,
          legend=TRUE, 
          legend.title="IAV Status",
          style="lattice",
          title="PLS-DA Comp 1 - 2")

# Set x and y axis labels based on model loadings
labelx <- paste("X-variate 1: ", 
                round(abs(plsda_model$prop_expl_var$X[1]*100), digits=0),
                "% expl. var",
                sep="")

labely <- paste("X-variate 2: ", 
                round(abs(plsda_model$prop_expl_var$X[2]*100), digits=0),
                "% expl. var",
                sep="")

# ggplot2
plsda_ggplot <- 
  data.frame(x = plsda_model$variates$X[,1], 
             y = plsda_model$variates$X[,2],
             iav = plsda_model$Y)

iav_plsda_ggplot <- 
  ggplot(plsda_ggplot, 
         aes(x=x, 
             y=y,
             col=iav)) +
  geom_point() +
  stat_ellipse() +
  labs(x=labelx,
       y = labely,
       name="IAV Status") +
  scale_color_manual("IAV Status",
                     values=c("#9d9d9d","#008ba2")) +
  theme_bw() +
  theme(panel.grid=element_blank(),
        legend.position=c(0.1, 0.1))

iav_plsda_ggplot
```

#### Cytokine Contributions
```{r}
plsda_loadings <- 
  plsda_model$loadings$X %>% 
  data.frame() %>% 
  select(1:2) %>% 
  arrange(comp1)
plsda_loadings

biplot(plsda_model, ind.names=FALSE, legend.title="IAV Status")

plotVar(plsda_model)
```


### 2-D Kolmogorov-Smirnov test
```{r}
ks_data <- 
  data.frame(plsda_model$variates$X,
             iav=plsda_model$Y)

plsda_pos <- 
  ks_data %>% 
  filter(iav=="IAV+") %>% 
  select(comp1, comp2)

plsda_neg <- 
  ks_data %>% 
  filter(iav=="IAV-") %>% 
  select(comp1, comp2)

fasano.franceschini.test(plsda_pos, plsda_neg)
```