---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# IAV ~ cytokines (2)
Analysis dates: 2016 and 2022

## Libraries
```{r, warning=FALSE, message=FALSE}
library(vegan)
library(mixOmics)
library(tidyverse)
library(pairwiseAdonis)
library(mlr3)
library(mlr3tuning)
library(rpart)
library(rpart.plot)
library(ggmosaic)
library(ggpubr)
library(EnvStats)
library(gridExtra)
library(grid)
```

## Data
```{r, message=FALSE, warning=FALSE}
cyto <- 
  read.csv("Output Files/cleaned_cytokine.csv") %>% 
  filter(analysis.year=="2016" | analysis.year=="2022")

cytobin <- 
  read.csv("Output Files/cleaned_cytokine_bin.csv") %>% 
  filter(analysis.year=="2016" | analysis.year=="2022")

meta <- read.csv("Output Files/metadata_cyto.csv") %>% 
  mutate(iav=gsub("neg", "IAV-", iav),
         iav=gsub("pos", "IAV+", iav))

# Merge cytokine & metadata 
data <- merge(cyto, meta, by="Sample") %>% 
  filter(!is.na(iav))

databin <- 
  merge(cytobin, meta, by="Sample") %>% 
  filter(!is.na(iav))
```


## PCA
```{r}
pca <- 
  prcomp(data[,2:10], center=TRUE, scale=TRUE)

pca_df <- 
  data.frame(
    x=pca$x[,1],
    y=pca$x[,2],
    molt.stage=data$molt.stage,
    year=factor(data$year),
    location=data$location,
    sex=data$sex,
    iav=data$iav,
    iavser=data$iavser,
    bc_bin=data$bc_bin,
    analysis.year=factor(data$analysis.year),
    batch=factor(data$batch)
    )

pca_labelx <-
  paste("PC1 (", 
        round(abs(summary(pca)$importance[2,1]*100), digits=0),
        "%)",
        sep="")
pca_labely <-
  paste("PC2 (", 
        round(abs(summary(pca)$importance[2,2]*100), digits=0),
        "%)",
        sep="")
```

### Analysis Year
Year that samples were analyzed, "batches"
```{r}
pca_analysisyear <- 
  ggplot(pca_df, 
         aes(x=x, 
             y=y,
             col=analysis.year)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
pca_analysisyear
```

### IAV
Year that samples were analyzed, "batches"
```{r}
pca_iav <- 
  ggplot(pca_df, 
         aes(x=x, 
             y=y,
             col=iav)) +
  geom_point() +
  stat_ellipse() +
  labs(x=pca_labelx,
       y = pca_labely) +
  theme_bw() +
  theme(panel.grid=element_blank())
pca_iav
```


## PERMANOVA

### Cytokine Presence/Absence

#### Create similarity matrix (Sorensen)
```{r}
databin_perm <- 
  databin %>% 
  select(2:10)

databin_dist <- 
  betadiver(databin_perm, method=11)

plot(databin_dist)

# Dissimilarity summary stats
mean(1-databin_dist); min(1-databin_dist); max(1-databin_dist); median(1-databin_dist)
```

#### Run PERMANOVA
```{r}
databin_adonis <- 
  adonis2((1-databin_dist) ~ iav, data=databin, permutations=4999)
databin_adonis

# R2
databin_adonis$R2[1]*100
```

#### Visualization
Principal Coordinates Analysis
```{r, warning=FALSE}
# Use 1 - Sorensen's for dissimilarity 
databin_pco <- cmdscale(1-databin_dist, eig=TRUE)

plot(databin_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)

ordispider(jitter(databin_pco$points, factor=1), factor(databin$iav), show.groups="IAV+", 
           col="deepskyblue", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(databin_pco$points, factor=1), factor(databin$iav), show.groups="IAV-", 
           col="firebrick", label = TRUE, display="sites", lwd=1.5, cex=1.2)
```


#### Homogeneity of group dispersions
```{r}
disper <- 
  betadisper(databin_dist, group=databin$iav, type="centroid")
anova(disper)
disper
boxplot(disper)
```


### Cytokine Concentration
#### Run PERMANOVA
```{r}
data_perm <-
  data %>% 
  select(2:10)

data_dist <- 
  vegdist(data_perm, method="bray")

data_adonis <- adonis2(data_dist ~ iav, data=data, 
                       permuations=4999)
data_adonis

# R2
data_adonis$R2[1]*100
```

#### Visualization
```{r, warning=FALSE}
# Use Bray-Curtis dissimilarity matrix
data_pco <- cmdscale(data_dist, eig=TRUE)

plot(data_pco$points, type="n", 
     cex.lab=1.5, cex.axis=1.1, cex.sub=1.1)

ordispider(jitter(data_pco$points, factor=1), factor(data$iav), show.groups="IAV+", 
           col="deepskyblue", label = TRUE, display="sites", lwd=1.5, cex=1.2)
ordispider(jitter(data_pco$points, factor=1), factor(data$iav), show.groups="IAV-", 
           col="firebrick", label = TRUE, display="sites", lwd=1.5, cex=1.2)
```

#### Homogeneity of group dispersions
```{r}
disper <- 
  betadisper(data_dist, group=data$iav, type="centroid")
anova(disper)
boxplot(disper)
```


## Cytokine Importance
### Prepare data
```{r, message=FALSE}
data_cart <- 
  data %>% 
  select(2:10, iav) %>% 
  mutate(iav=factor(iav))
```

### Cross Validation
```{r}
# Creating task and learner
task <- as_task_classif(iav ~ ., 
                        data = data_cart)
        
task <- task$set_col_roles(cols="iav",
                           add_to="stratum")

learner <- lrn("classif.rpart",
               predict_type = "prob",
               maxdepth = to_tune(2, 5),
               minbucket = to_tune(1, 39),
               minsplit = to_tune(1, 30)
               )

# Define tuning instance - info ~ tuning process
instance <- ti(task = task,
               learner = learner,
               resampling = rsmp("cv", folds = 10),
               measures = msr("classif.ce"),
               terminator = trm("none")
               )

# Define how to tune the model
tuner <- tnr("grid_search", 
             batch_size = 10
             )

# Trigger the tuning process
#tuner$optimize(instance)

# optimal values: 
  # maxdepth = 3
  # minbucket = 9
  # minsplit = 20
  # classif.ce = 0.3230556
```

### Run model with optimized parameters
```{r}
cart_model <- rpart(formula=iav ~ .,
                    data=data_cart,
                    method="class",
                    maxdepth=3,
                    minbucket=9,
                    minsplit=20)

# plot optimized tree
rpart.plot(cart_model)
cart_model$variable.importance 
  # IL-10, IL-18
```


### Plot
Variable importance: the sum of the goodness of split measures for each split for which it was the primary variable, plus goodness * (adjusted agreement) for all splits in which it was a surrogate. In the printout these are scaled to sum to 100 and the rounded values are shown, omitting any variable whose proportion is less than 1%. (https://cran.r-project.org/web/packages/rpart/vignettes/longintro.pdf)

#### Variable Importance
```{r}
varimp <- 
  data.frame(imp=cart_model$variable.importance) %>% 
  rownames_to_column() %>% 
  rename("variable" = rowname) %>% 
  arrange(imp) %>%
  mutate(variable=gsub("\\.","-",variable),
         variable = forcats::fct_inorder(variable))

varimp_plot <- 
  ggplot(varimp) + 
  geom_segment(aes(x = variable, 
                   y = 0, 
                   xend = variable, 
                   yend = imp), 
               linewidth = 0.5, 
               alpha = 0.7) +
  geom_point(aes(x = variable, 
                 y = imp), 
             size = 1, 
             show.legend = F,
             col="black") +
  labs(x="Cytokine", 
       y="Variable Importance") +
  coord_flip() +
  theme_classic() +
  theme(text=element_text(size=10, color="black"))
```

#### Data Prep
```{r}
databin <-
  databin %>% 
  mutate(across(2:10, \(x) gsub("1", "Present", x)),
         across(2:10, \(x) gsub("0", "Absent", x)),
         across(2:10, \(x) factor(x, levels=c("Present", "Absent"))))
```

#### Proportions
```{r}
prop.table(table(databin$iav, databin$IL.10), margin=1)*100
prop.table(table(databin$iav, databin$IL.18), margin=1)*100

il10_prop <- 
  databin %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.10,iav), fill=IL.10)) +
  scale_fill_manual(values=c("Present"="gray60", "Absent"="lightgray"),
                    name="Cytokine") +
  scale_y_continuous(labels = scales::percent) +
  labs(title="IL-10") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position="bottom",
        text = element_text(size=10),
        plot.title = element_text(hjust = 0.5))
il10_prop

cyto_leg <- get_legend(il10_prop)

il10_prop <-
  il10_prop +
  theme(legend.position="none")

il18_prop <- 
  databin %>%
  ggplot(data=.) +
  geom_mosaic(aes(x=product(IL.18,iav), fill=IL.18)) +
  scale_fill_manual(values=c("Present"="gray60", "Absent"="lightgray")) +
  scale_y_continuous(labels = scales::percent) +
  labs(title="IL-18") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.position="none",
        text = element_text(size=10),
        plot.title = element_text(hjust = 0.5))
il18_prop
```


#### Concentrations
```{r}
il10 <- 
  data %>% 
  select(iav, IL.10) %>% 
  filter(!IL.10==0) %>% 
  ggplot(., aes(x=iav, y=log(IL.10))) +
  geom_boxplot(color="black", fill="gray60", lwd=0.25,
               outliers=FALSE) +
  geom_point(size=0.75, position=position_jitter(width=0.2)) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=3) +
  theme_classic() +
  theme(text = element_text(size=10),
         panel.border=element_rect(color="black", fill=NA, linewidth=0.5))
il10

il18 <- 
  data %>% 
  select(iav, IL.18) %>% 
  filter(!IL.18 == 0) %>% 
  ggplot(., aes(x=iav, y=log(IL.18))) +
  geom_boxplot(color="black", fill="gray60", lwd=0.25,
               outliers=FALSE) +
  geom_point(size=0.75, position=position_jitter(width=0.2)) +
  labs(x=NULL, y=NULL) +
  stat_n_text(size=3) +
  theme_classic() +
  theme(text = element_text(size=10),
        panel.border=element_rect(color="black", fill=NA, linewidth=0.5))
il18
```

#### Combine together
```{r}
prop_grid <- grid.arrange(il10_prop, il18_prop, ncol=2,
                          left=textGrob("Proportion of Samples",
                                        rot=90, gp=gpar(fontsize=10)))
prop_grid <- grid.arrange(prop_grid, cyto_leg,
                          ncol=1, 
                          heights=c(3, 0.3))
conc_grid <- grid.arrange(il10, il18, ncol=2, 
                          left=textGrob("log(Concentration (pg/mL))", 
                                        rot=90, gp=gpar(fontsize=10)))

cyto_grid <- grid.arrange(prop_grid, conc_grid,
                          ncol=1, 
                          heights=c(4,3.5),
                          bottom=textGrob("IAV Status", gp=gpar(fontsize=10)))

final_grid <- 
  grid.arrange(varimp_plot, cyto_grid, 
               ncol=1, heights=c(2,4)) 
```


## How different IAV+ and IAV- groups are

### PLSDA 

#### Prepare data
```{r}
pls_x <-
 data_cart[,1:9] 

pls_y <- 
  data_cart %>% 
  mutate(iav=gsub("neg", "IAV-", iav),
         iav=gsub("pos", "IAV+", iav),
         iav=as.factor(iav)) %>% 
  select(iav)
```

#### PLSDA Model
```{r}
plsda_model <- mixOmics::plsda(pls_x,
                               pls_y$iav,
                               scale=TRUE,
                               ncomp=9)
```

#### Visualizations
```{r}
# Set x and y axis labels based on model loadings
labelx <- 
  paste("X-variate 1: ", 
        round(abs(plsda_model$prop_expl_var$X[1]*100), digits=0),
        "% expl. var",
        sep="")

labely <- 
  paste("X-variate 2: ", 
        round(abs(plsda_model$prop_expl_var$X[2]*100), digits=0),
        "% expl. var",
        sep="")

# ggplot2
plsda_ggplot <- 
  data.frame(x = plsda_model$variates$X[,1], 
             y = plsda_model$variates$X[,2],
             iav = plsda_model$Y,
             analysis.year=factor(data$analysis.year))

iav_plsda_ggplot <- 
  ggplot(plsda_ggplot, 
         aes(x=x, 
             y=y,
             col=iav)) +
  geom_point() +
  stat_ellipse() +
  labs(x=labelx,
       y = labely,
       name="IAV Status") +
  scale_color_manual("IAV Status",
                     values=c("#9d9d9d","#008ba2")) +
  theme_bw() +
  theme(panel.grid=element_blank())

iav_plsda_ggplot
```

#### Cytokine Contributions
```{r}
plsda_loadings <- 
  plsda_model$loadings$X %>% 
  data.frame() %>% 
  select(1:2) %>% 
  arrange(comp1)
plsda_loadings

biplot(plsda_model, ind.names=FALSE, legend.title="IAV Status")

plotVar(plsda_model)
```
