---
author: "Christina McCosker"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# IAV ~ swabs

## Load required libraries
```{r, warning=FALSE, message=FALSE}
library(mixOmics)
library(tidyverse)
library(ggfortify)
library(gridExtra)
library(cramer)
```

## Read in data 
```{r, message=FALSE, warning=FALSE}
# Cytokines
cyto <- 
  read.csv("Output Files/cleaned_cytokine.csv")
cytobin <- 
  read.csv("Output Files/cleaned_cytokine_bin.csv")
cytoscale <- 
  read.csv("Output Files/cleaned_cytokine_scaled.csv")

# Metadata
meta <- 
  read.csv("Output Files/metadata_cyto.csv")

# Swab/virology data
viro <- 
  read.csv("Input Files/Hg_virology_2023.csv", sep=',', strip.white=TRUE) %>% 
  select(Tag.ID., year, IAV, IAV.Nasal, IAV.Conj, IAV.Rectal) %>% 
  filter(Tag.ID. %in% meta$Sample)
```

## Tidy & Merge Data
```{r}
# Clean up swab data
viro_tidy <-
  viro %>% 
  mutate(IAV.Nasal=gsub(">45|Undetermined","neg", IAV.Nasal),
         IAV.Nasal=gsub("Pooled\\(POS\\)", "POS", IAV.Nasal),
         IAV.Nasal=gsub("\\#N\\/A", NA, IAV.Nasal),
         IAV.Nasal=ifelse(IAV.Nasal < 45, "POS", IAV.Nasal),
         IAV.Rectal=gsub(">45|Undetermined","neg", IAV.Rectal),
         IAV.Rectal=gsub("Pooled\\(POS\\)", "POS", IAV.Rectal),
         IAV.Rectal=gsub("ND|\\#N\\/A", NA, IAV.Rectal),
         IAV.Rectal=ifelse(IAV.Rectal < 45, "POS", IAV.Rectal),
         IAV.Conj=gsub(">45|Undetermined","neg", IAV.Conj),
         IAV.Conj=gsub("\\#N\\/A", NA, IAV.Conj),
         IAV.Conj=gsub("Pooled\\(POS\\)", "POS", IAV.Conj),
         IAV.Conj=ifelse(IAV.Conj < 45, "POS", IAV.Conj))


# Merge cytokine, meta, virology data
data <-
  cyto %>% 
  merge(., viro_tidy, by.x="Sample", by.y="Tag.ID.") %>% 
  mutate(IAV=meta$iav)

databin <-
  cytobin %>% 
  merge(., viro_tidy, by.x="Sample", by.y="Tag.ID.") %>% 
  mutate(IAV=meta$IAV)

datascale <- 
  cytoscale %>% 
  merge(., viro_tidy, by.x="Sample", by.y="Tag.ID.") %>% 
  mutate(IAV=meta$iav)
```

## PCA on log-transformed, scaled, centered cytokine data
```{r}
# Conj
data_conj <- 
  datascale %>% 
  filter(!is.na(IAV.Conj))

pca_conj <- 
  prcomp(data_conj[,2:10], center=FALSE, scale=FALSE)

plot_pca_conj <- 
  autoplot(pca_conj, data=data_conj, shape="IAV.Conj") + 
  scale_shape_manual(values=c(1, 19), 
                     name="Conj Swab") +
  theme_bw() +
  theme(panel.grid=element_blank(), text=element_text(size=8),
        legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
        legend.position="bottom")

# Nasal
data_nasal <- 
  datascale %>% 
  filter(!is.na(IAV.Nasal))

pca_nasal <- 
  prcomp(data_nasal[,2:10], center=FALSE, scale=FALSE)

plot_pca_nasal <- 
  autoplot(pca_nasal, data=data_nasal, shape="IAV.Nasal") + 
  scale_shape_manual(values=c(1, 19), 
                     name="Nasal Swab") +
  theme_bw() +
  theme(panel.grid=element_blank(), text=element_text(size=8),
        legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
        legend.position="bottom")

# Rectal
data_rectal <- 
  datascale %>% 
  filter(!is.na(IAV.Rectal))

pca_rectal <- 
  prcomp(data_rectal[,2:10], center=FALSE, scale=FALSE)

plot_pca_rectal <- 
  autoplot(pca_rectal, data=data_rectal, shape="IAV.Rectal") + 
  scale_shape_manual(values=c(1, 19), 
                     name="Rectal Swab") +
  theme_bw() +
  theme(panel.grid=element_blank(), text=element_text(size=8),
        legend.background=element_rect(size=0.5, linetype="solid", colour ="black"),
        legend.position="bottom")

grid.arrange(arrangeGrob(plot_pca_conj, plot_pca_nasal, plot_pca_rectal, ncol=3))

```

## Test for normality
```{r}
# Create dataframe for results
norm_results <- data.frame(matrix(ncol=0, nrow=9)) %>%
  mutate(cyto=colnames(datascale)[2:10],
         W=NA, 
         pvalue=NA)

# Truncate data
data_norm <-
  datascale[,2:10]

# Test!
for (i in 1:9) {
  test <- shapiro.test(data_norm[,i])
  norm_results[i,2] <- round(test$statistic, digits=3)
  norm_results[i,3] <- test$p.value
}

norm_results # all not normally distributed
```

## Cramer Test 
Multivariate composition of the cytokine pattern between each infection stage group with its control. It can handle multiple variables, potentially complex interactions and correlations.  
```{r}
# Conj Swabs
con_pos <-
  datascale %>% 
  filter(IAV.Conj=="POS") %>% 
  select(1:9) %>% 
  data.matrix()
  
con_neg <- 
  datascale %>% 
  filter(IAV.Conj=="neg") %>% 
  select(1:9) %>% 
  data.matrix() 

cramer.test(con_pos, con_neg)


# Nasal Swabs
nas_pos <-
  datascale %>% 
  filter(IAV.Nasal=="POS") %>% 
  select(1:9) %>% 
  data.matrix()
  
nas_neg <- 
  datascale %>% 
  filter(IAV.Nasal=="neg") %>% 
  select(1:9) %>% 
  data.matrix() 

cramer.test(nas_pos, nas_neg)


# Rectal Swabs
rec_pos <-
  datascale %>% 
  filter(IAV.Rectal=="POS") %>% 
  select(1:9) %>% 
  data.matrix()
  
rec_neg <- 
  datascale %>% 
  filter(IAV.Rectal=="neg") %>% 
  select(1:9) %>% 
  data.matrix() 

cramer.test(rec_pos, rec_neg)

# Between positive swabs?
cramer.test(con_pos, nas_pos) # early vs mid infection
cramer.test(con_pos, rec_pos) # early vs late infection
cramer.test(nas_pos, rec_pos) # mid vs late infection
```

## Cytokine Exploratory Plots ~ Swabs
```{r}
# Configure data
swabs <-
  data %>% 
  select(!year & !IAV) %>% 
  pivot_longer(cols=c(2:10), names_to="cytokine", values_to="concentration") %>% 
  pivot_longer(cols=c(2:4), names_to="swab", values_to="status") 

conj <- 
  swabs %>% 
  filter(swab=="IAV.Conj", 
         status=="POS")
  
nasal <- 
  swabs %>% 
  filter(swab=="IAV.Nasal",
         status=="POS") 

rectal <- 
  swabs %>% 
  filter(swab=="IAV.Rectal", 
         status=="POS")

pos_swabs <-
  rbind(conj, nasal, rectal)

# Plots
ifn <-
  pos_swabs %>% 
  filter(cytokine=="IFNg") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IFNg")

# Plots
il2 <- 
  pos_swabs %>% 
  filter(cytokine=="IL.2") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IL.2")

il6 <- 
  pos_swabs %>% 
  filter(cytokine=="IL.6") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IL.6")

il7 <- 
  pos_swabs %>% 
  filter(cytokine=="IL.7") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IL.7")

il8 <- 
  pos_swabs %>% 
  filter(cytokine=="IL.8") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IL.8")

il10 <- 
  pos_swabs %>% 
  filter(cytokine=="IL.10") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IL.10")

il15 <- 
  pos_swabs %>% 
  filter(cytokine=="IL.15") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IL.15")

il18 <- 
  pos_swabs %>% 
  filter(cytokine=="IL.18") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="IL.18")

kc <- 
  pos_swabs %>% 
  filter(cytokine=="KC.like") %>% 
  ggplot(., aes(x=swab, y=log(concentration))) +
  geom_boxplot() +
  labs(x=NULL, y="KC.like")

imp <- grid.arrange(arrangeGrob(il6, il10, kc))
nonimp1 <- grid.arrange(arrangeGrob(ifn, il2, il7))
nonimp2 <- grid.arrange(arrangeGrob(il8, il15, il18))

all <- grid.arrange(imp, nonimp1, nonimp2, ncol=3)

ggsave(filename="Figures/cyto_swab.jpeg", all, width=8, height=6, units="in")
```